---
title: "Module 00 — Environment setup and readiness checks"
author: "HWC North Africa automated evidence map"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, dpi = 300)
library(here)
source(here::here("R", "config.R"))
```

# Purpose

Module 00 does exactly two things: scaffold the project and prove that each
external dependency the pipeline needs can actually be reached from this
machine. Anything that fails here must be resolved before Module 01. **No
downstream module is permitted to substitute mock output for a failed
dependency.**

# Directory scaffold

```{r scaffold-check}
expected <- c(
  DIR_RAW, DIR_DERIVED, DIR_EXTERNAL, DIR_BENCHMARKS,
  DIR_FIGURES, DIR_TABLES, DIR_OUTPUT, DIR_LOGS,
  "R", "rmd"
)
present <- vapply(expected, function(p) dir.exists(here::here(p)), logical(1))
stopifnot(all(present))
data.frame(path = expected, exists = present)
```

# R toolchain

```{r r-toolchain}
r_ver  <- R.version.string
r_plat <- R.version$platform
have_renv <- requireNamespace("renv", quietly = TRUE)

data.frame(
  item  = c("R version", "Platform", "renv installed"),
  value = c(r_ver, r_plat, as.character(have_renv))
)
```

# Required R packages

The pipeline expects these packages. Missing ones are listed but **not
auto-installed** — installation is an explicit step so the user can review it.

```{r r-packages}
required <- c(
  "openalexR", "rcrossref", "httr2", "jsonlite", "xml2",
  "dplyr", "tidyr", "readr", "stringr", "stringi", "purrr",
  "here", "fs", "digest", "glue",
  "ggplot2", "scales", "viridisLite", "patchwork",
  "knitr", "rmarkdown", "sessioninfo"
)
have <- vapply(required, function(p) requireNamespace(p, quietly = TRUE), logical(1))
missing_pkgs <- required[!have]
data.frame(package = required, installed = have)
if (length(missing_pkgs) > 0) {
  message("Missing packages: ", paste(missing_pkgs, collapse = ", "))
}
```

# OpenAlex reachability (live — no key required)

The protocol names two anchor findings we should be able to retrieve from a
North Africa search: wild boar work from the Ghandri group in Tunisia, and
Gulf of Gabès turtle bycatch. If these do not surface, either the search
grammar is wrong or the endpoint is unreachable.

```{r openalex-anchor, eval=requireNamespace("openalexR", quietly=TRUE)}
library(openalexR)

boar_hits <- oa_fetch(
  entity      = "works",
  search      = "wild boar Tunisia Sus scrofa damage",
  per_page    = 25,
  verbose     = FALSE
)

turtle_hits <- oa_fetch(
  entity      = "works",
  search      = "Gulf of Gabes turtle bycatch Caretta",
  per_page    = 25,
  verbose     = FALSE
)

data.frame(
  probe    = c("Tunisia wild boar", "Gulf of Gabès turtle bycatch"),
  n_hits   = c(NROW(boar_hits), NROW(turtle_hits)),
  earliest = c(min(boar_hits$publication_year,  na.rm = TRUE),
               min(turtle_hits$publication_year, na.rm = TRUE)),
  latest   = c(max(boar_hits$publication_year,  na.rm = TRUE),
               max(turtle_hits$publication_year, na.rm = TRUE))
)
```

# CORE API key

```{r core-key}
key_set <- !is.na(CORE_API_KEY) && nzchar(CORE_API_KEY)
data.frame(check = "CORE_API_KEY present", value = key_set)
```

If `FALSE`, register for a free key at <https://core.ac.uk/services/api> and add
to `~/.Renviron` as `CORE_API_KEY=...`, then restart R.

# GROBID reachability

```{r grobid-check}
grobid_ok <- tryCatch({
  r <- httr::GET(paste0(GROBID_URL, "/api/isalive"), httr::timeout(3))
  httr::status_code(r) == 200
}, error = function(e) FALSE)
data.frame(check = "GROBID /api/isalive", value = grobid_ok)
```

If `FALSE`: install Docker Desktop, start it, then run
`docker run --rm -p 8070:8070 lfoppiano/grobid:0.8.1` (ARM64 image supported).

# Ollama model server reachability + ensemble present

```{r ollama-check}
ollama_ok <- tryCatch({
  r <- httr::GET(paste0(OLLAMA_URL, "/api/tags"), httr::timeout(3))
  httr::status_code(r) == 200
}, error = function(e) FALSE)

installed_models <- character(0)
if (ollama_ok) {
  tags <- jsonlite::fromJSON(paste0(OLLAMA_URL, "/api/tags"))
  installed_models <- tags$models$name
}
ensemble_present <- ENSEMBLE_MODELS %in% installed_models

data.frame(
  check = c("Ollama /api/tags reachable",
            paste("model present:", ENSEMBLE_MODELS)),
  value = c(ollama_ok, ensemble_present)
)
```

If any model is missing: `ollama pull <name>`. At 16 GB unified memory, serve
the ensemble models **sequentially**, not concurrently — Ollama unloads on
its own between calls but you can force it with `ollama stop <model>`.

# Ensemble JSON schema probe

Each model must return schema-valid JSON at temperature 0 before it is
allowed to classify anything. Failing this check disqualifies the model.

```{r ollama-json-probe, eval=ollama_ok}
probe_one <- function(model) {
  body <- list(
    model  = model,
    prompt = "Return a JSON object with fields include (true|false), confidence (low|medium|high), reason (string). Item: \"Population dynamics of wild boar (Sus scrofa) in northern Tunisia and their impact on cereal crops\".",
    format = "json",
    stream = FALSE,
    options = list(temperature = OLLAMA_TEMPERATURE, seed = OLLAMA_SEED)
  )
  r <- httr::POST(paste0(OLLAMA_URL, "/api/generate"),
                  body = jsonlite::toJSON(body, auto_unbox = TRUE),
                  encode = "raw",
                  httr::content_type_json(),
                  httr::timeout(120))
  raw <- httr::content(r, "parsed", "application/json")$response
  parsed <- tryCatch(jsonlite::fromJSON(raw), error = function(e) NULL)
  list(model = model,
       ok    = !is.null(parsed) && all(c("include","confidence","reason") %in% names(parsed)),
       raw   = substr(raw, 1, 240))
}
probes <- lapply(ENSEMBLE_MODELS, probe_one)
data.frame(
  model         = vapply(probes, `[[`, "", "model"),
  schema_valid  = vapply(probes, `[[`, logical(1), "ok"),
  first_240ch   = vapply(probes, `[[`, "", "raw")
)
```

# Note on Unpaywall (removed from stack)

The protocol originally included Unpaywall for OA-status lookup. Removed
because OpenAlex already ingests Unpaywall's OA metadata on every work
record (`best_oa_location.pdf_url`), so a direct Unpaywall dependency was
redundant and would have added a ToS-mandated contact-email query parameter
for no additional coverage. CORE remains the primary fulltext source and
provides its own OA download URLs.

# Session info

```{r sessioninfo}
sessionInfo()
```
