---
title: "Module 01 — Eligibility protocol (written and frozen before searching)"
author: "HWC North Africa automated evidence map"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, dpi = 300)
library(here)
source(here::here("R", "config.R"))
source(here::here("R", "eligibility.R"))
```

# What this document is, and why it is committed before Module 02

An evidence map's defensibility rests on the fact that its inclusion rules
were **written before the corpus existed**. Any rule invented mid-screening
is a rule tuned to the data. This document is the pre-registration: it is
committed to git before `02_search.Rmd` is executed, and any subsequent edit
is a dated amendment with a stated reason.

Every rule below is also present in `R/eligibility.R`, a machine-readable
copy that Module 04b sources directly into the classifier prompts. There is
one source of truth; the two files cannot drift.

```{r design-statement, results='asis', echo=FALSE}
cat("> **Design:** ", ELIGIBILITY$design, "\n", sep = "")
```

# PCC framework

We use JBI's Population–Concept–Context frame, which is the standard for
scoping-family reviews and evidence maps.

```{r pcc, echo=FALSE}
pcc <- ELIGIBILITY$pcc
data.frame(
  element = c("Population", "Concept", "Context (countries)", "Context (realms)"),
  content = c(
    pcc$population,
    paste(pcc$concept, collapse = ", "),
    paste(pcc$context$countries, collapse = ", "),
    paste0("terrestrial = ", pcc$context$terrestrial,
           "; marine = ",     pcc$context$marine)
  )
)
```

# Inclusion criteria

A record must satisfy **every** clause to be included.

```{r include, echo=FALSE}
knitr::kable(data.frame(`#` = seq_along(ELIGIBILITY$include),
                        clause = ELIGIBILITY$include),
             check.names = FALSE)
```

# Exclusion criteria

A record is excluded if **any** clause applies.

```{r exclude, echo=FALSE}
knitr::kable(data.frame(`#` = seq_along(ELIGIBILITY$exclude),
                        clause = ELIGIBILITY$exclude),
             check.names = FALSE)
```

# Borderline case rules

These are the six cases the protocol names as ambiguous. Each is decided
here, not during classification, so that ambiguity converts into
consistent-but-imperfect noise instead of drifting rule application.

The verdicts are:

```{r borderline, echo=FALSE, results='asis'}
for (nm in names(ELIGIBILITY$borderline)) {
  b <- ELIGIBILITY$borderline[[nm]]
  cat("## ", gsub("_", " ", nm), "\n\n", sep = "")
  cat("**Verdict:** ", b$verdict, "\n\n", sep = "")
  cat("**Rule:** ", b$rule, "\n\n", sep = "")
  cat("**Rationale:** ", b$rationale, "\n\n---\n\n", sep = "")
}
```

# Language, date, geography

```{r bounds, echo=FALSE}
data.frame(
  parameter = c("Earliest publication year",
                "Languages searched (no machine translation)",
                "Countries",
                "Realms"),
  value = c(
    as.character(ELIGIBILITY$year_from),
    paste(ELIGIBILITY$languages, collapse = ", "),
    paste(ELIGIBILITY$pcc$context$countries, collapse = ", "),
    paste0("terrestrial + marine (marine kept because Mediterranean bycatch is a documented HWC issue for Gulf of Gabès sea turtles and Mediterranean monk seals)")
  )
)
```

Arabic-language search yield is expected to be low. **Low yield is an
indexing finding, not evidence of absent research** — it will be reported as
such rather than treated as an empirical count.

# Grey literature

```{r grey, echo=FALSE, results='asis'}
g <- ELIGIBILITY$grey_literature
cat("- **Included:** ", g$include, "\n", sep = "")
cat("- **Stratum rule:** ", g$stratum_rule, "\n", sep = "")
```

# Traceability into downstream modules

The eligibility list in `R/eligibility.R` is consumed unmodified by:

- **`04b_classify.Rmd`** — the six borderline rules are interpolated
  verbatim into the classification prompt for both ensemble models.
- **`06_extract.Rmd`** — the inclusion/exclusion clauses become the
  "eligibility_reason" enum for the primary_reason field.
- **`09_outputs.Rmd`** — the design statement appears verbatim in the
  methods paragraph.

There is no re-typing, and no free-text version of the criteria maintained
anywhere else in the codebase. Grepping the repo for a criterion phrase
must return exactly one authoring location.

# Amendments log

Amendments to the protocol after this document is committed must appear as
dated entries below, each stating the change, the reason, and the commit
hash where it landed. Silent edits are a protocol violation.

- **2026-08-11** — Removed direct Unpaywall dependency from the stack.
  Reason: OpenAlex ingests Unpaywall OA data on every work record
  (`best_oa_location.pdf_url`), making the direct dependency redundant.
  Commit: `84feb99`.
- **2026-08-11** — Substituted `mlx_lm.server` (planned) with Ollama
  (actually available on the target machine). Ensemble is now
  `gemma4:e4b` (Gemma family) plus `llama3:latest` (Llama family), served
  sequentially at 16 GB unified memory. Reason: match the actual local
  toolchain; preserve the "genuinely different training lineages"
  requirement. Commit: `84feb99`.

# Session info

```{r sessioninfo}
sessionInfo()
```
