---
title: "Module 02 — Search execution and ledger"
author: "HWC North Africa automated evidence map"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, dpi = 300)
suppressPackageStartupMessages({
  library(here); library(dplyr); library(tibble); library(readr)
})
source(here::here("R", "config.R"))
source(here::here("R", "eligibility.R"))
source(here::here("R", "search_terms.R"))
```

# What ran, and when

The full search matrix is executed by `R/run_searches.R` and cached at
`data/raw/search/*.rds`. This knit **only reads the cache** — it does not
issue API requests. If a cell file is missing, the corresponding row will
report zero and the reader will know to re-run the runner.

```{r read-ledger}
ledger_path <- here::here(DIR_DERIVED, "search_ledger.rds")
stopifnot("search_ledger.rds not found — run Rscript R/run_searches.R first" =
          file.exists(ledger_path))
ledger <- readRDS(ledger_path)
cat("Ledger rows (raw, pre-dedup):", NROW(ledger), "\n")
cat("Ledger written:              ", format(file.info(ledger_path)$mtime), "\n")
```

# Search cells and their sizes

Cell design: per-country English searches on each source; one pooled French
cell per source (French geography terms cover all 5 countries); one pooled
Arabic cell per source with the transliteration decisions documented in
`R/search_terms.R`.

```{r per-cell}
per_cell <- ledger |>
  dplyr::count(cell_id, source, name = "n_records") |>
  dplyr::arrange(cell_id)
knitr::kable(per_cell)
```

# Source × language cross-tab

```{r source-lang}
lang_from_cell <- function(x) sub("^[a-z]+_(.*?)_.*", "\\1", x)
country_from_cell <- function(x) sub("^[a-z]+_[a-z]{2}_", "", x)
ledger_tab <- ledger |>
  dplyr::mutate(lang = lang_from_cell(cell_id),
                country_code = country_from_cell(cell_id))
xt <- ledger_tab |>
  dplyr::count(source, lang, name = "n") |>
  tidyr::pivot_wider(names_from = lang, values_from = n, values_fill = 0L)
knitr::kable(xt, caption = "Records per source × language (pre-dedup)")
```

# Coverage sanity — the two anchor findings

The protocol names two known-good North Africa results. Both must appear in
the ledger. If they don't, either the search grammar is broken or the
databases are unreachable.

```{r anchors}
anchor_boar <- ledger |>
  dplyr::filter(grepl("Tunisia", title, ignore.case = TRUE),
                grepl("boar|Sus scrofa", title, ignore.case = TRUE))
anchor_turtle <- ledger |>
  dplyr::filter(grepl("Gabes|Gabès|Mediterran", title, ignore.case = TRUE),
                grepl("turtle|Caretta|bycatch|by-catch", title,
                      ignore.case = TRUE))
data.frame(
  anchor      = c("Tunisia + wild boar (in title)",
                  "Gulf of Gabès / Mediterranean turtle bycatch (in title)"),
  n_matches   = c(NROW(anchor_boar), NROW(anchor_turtle))
)
```

```{r anchors-top, results='asis'}
cat("### Sample boar-in-Tunisia matches (first 3):\n\n")
knitr::kable(head(anchor_boar |>
                  dplyr::select(publication_year, source, title), 3))
cat("\n### Sample turtle-bycatch matches (first 3):\n\n")
knitr::kable(head(anchor_turtle |>
                  dplyr::select(publication_year, source, title), 3))
```

# What is *not* in the ledger, and why

- **Crossref** — omitted 2026-08-12 by user direction. Rationale: Crossref
  is largely a DOI-authoritative subset of what OpenAlex already returns
  after DOI matching, and the marginal recall gain does not justify the
  extra rate-limit budget for a scoping study.
- **BASE / OpenAIRE (grey literature)** — deferred. Grey literature will
  be added as a separate stratum in a later commit; the protocol requires
  it be kept out of headline counts anyway.
- **Machine translation of English queries into French/Arabic** — never
  done, per protocol. Both non-English cells use human-selected terms.

# Query provenance

Every search cell's exact query, timing, and record count is appended to
`logs/02_search_YYYYMMDD.jsonl`. This is how the methods paragraph in
Module 09 will state the search strings verbatim without any risk of drift.

```{r log-tail}
today <- format(Sys.Date(), "%Y%m%d")
log_file <- here::here(DIR_LOGS, sprintf("02_search_%s.jsonl", today))
if (file.exists(log_file)) {
  cat("Log:", log_file, "\n")
  cat("Line count:", length(readLines(log_file)), "\n")
} else {
  cat("No log file for today at:", log_file, "\n")
}
```

# Session info

```{r sessioninfo}
sessionInfo()
```
