---
title: "Module 03 — Deduplication (DOI exact + title/year Jaro-Winkler)"
author: "HWC North Africa automated evidence map"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, dpi = 300)
suppressPackageStartupMessages({
  library(here); library(dplyr); library(tibble); library(jsonlite)
})
source(here::here("R", "config.R"))
```

# What this module does

Two passes, deterministic and reversible via the provenance columns.

1. **PASS 1 — DOI exact match on normalised DOI.** Same DOI → same work.
   Normalisation strips URL prefixes, lowercases, trims. Every merged
   record's `provenance_cells` and `provenance_sources` list-columns keep
   a full history of where each copy came from — nothing is lost.
2. **PASS 2 — Jaro-Winkler ≥ 0.95 on normalised titles, blocked by
   publication year ±1.** Applied only to the residue of records without
   a DOI. Titles are ASCII-folded, lowercased, punctuation stripped,
   leading articles dropped. Union-find within each year window.

Merge policy: when a group has multiple records, the OpenAlex record is
kept as canonical (richer metadata: abstract, is_oa, pdf_url,
relevance_score); the CORE record is kept only when there is no OpenAlex
alternative. Both cases preserve full provenance via the two list-columns.

Every merge and every pass boundary is logged to `logs/03_dedup.jsonl`
so the record arithmetic for the Module 08 flow diagram reconciles.

```{r read}
raw_path   <- here::here(DIR_DERIVED, "search_ledger.rds")
dedup_path <- here::here(DIR_DERIVED, "dedup_ledger.rds")
stopifnot(file.exists(raw_path), file.exists(dedup_path))
raw    <- readRDS(raw_path)
dedup  <- readRDS(dedup_path)
```

# Record arithmetic

```{r arithmetic}
tibble::tibble(
  stage           = c("Raw ledger (Module 02 output)",
                      "After Pass 1 (DOI exact)",
                      "After Pass 2 (title + year)",
                      "Deduplicated ledger (Module 03 output)"),
  n               = c(NROW(raw),
                      NA,   # filled in from log below
                      NA,
                      NROW(dedup)),
  n_collapsed_here = c(0,
                       NA, NA,
                       NROW(raw) - NROW(dedup))
) |> knitr::kable(caption = "Record loss ledger — every drop is a merge, not a deletion")
```

# Merge log summary

```{r merge-log}
log_path <- here::here(DIR_LOGS, "03_dedup.jsonl")
if (file.exists(log_path)) {
  lines <- readLines(log_path)
  entries <- lapply(lines, jsonlite::fromJSON)
  df <- dplyr::bind_rows(lapply(entries, as.data.frame))
  knitr::kable(df)
} else {
  cat("No dedup log at:", log_path, "\n")
}
```

# Source overlap

Records that appeared in more than one source are the strongest evidence
that CORE and OpenAlex are indexing genuinely different subsets.

```{r overlap}
overlap <- dedup |>
  dplyr::mutate(n_sources = vapply(provenance_sources, length, integer(1))) |>
  dplyr::count(n_sources, name = "n_records")
knitr::kable(overlap, caption = "Sources per deduplicated record")
```

# Cell overlap distribution

```{r cell-overlap}
cell_ovl <- dedup |>
  dplyr::mutate(n_cells = vapply(provenance_cells, length, integer(1))) |>
  dplyr::count(n_cells, name = "n_records") |>
  dplyr::arrange(dplyr::desc(n_records))
knitr::kable(head(cell_ovl, 15),
             caption = "How many search cells each unique record appeared in")
```

# Country attribution (from cell provenance)

Deduped records carry the union of every cell they appeared in, so a
paper that surfaced under both `openalex_en_TN` and `openalex_en_REGION`
is credited to Tunisia. Records appearing in `_REGION` only are shelved
to a "multi-country / regional" bucket for the map.

```{r country-attribution}
country_from_cells <- function(cells) {
  codes <- unique(sub("^[a-z]+_[a-z]{2}_", "", unlist(cells)))
  codes
}
dedup <- dedup |>
  dplyr::mutate(
    country_codes = lapply(provenance_cells, country_from_cells),
    n_countries   = vapply(country_codes, length, integer(1))
  )
country_long <- dedup |>
  dplyr::select(country_codes, n_merged) |>
  tidyr::unnest(country_codes) |>
  dplyr::count(country_codes, name = "n_records") |>
  dplyr::arrange(dplyr::desc(n_records))
knitr::kable(country_long,
             caption = "Records attributed to each country (union across cells)")
```

# Session info

```{r sessioninfo}
sessionInfo()
```
