---
title: "Module 04a — Ensemble classifier validation on SYNERGY"
author: "HWC North Africa automated evidence map"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, dpi = 300)
suppressPackageStartupMessages({
  library(here); library(dplyr); library(tibble); library(readr); library(ggplot2)
})
source(here::here("R", "config.R"))
```

# Why this module is load-bearing

The whole automated evidence map depends on one measured claim:
**how well the ensemble discriminates relevant from irrelevant records
when it is told the criteria it should apply.** Without a benchmark
number, "the classifier said so" is not evidence. With one, downstream
counts inherit a known error profile.

This module runs the *exact* ensemble, prompt template shape, and
decision rule the real corpus classification (Module 04b) will use, on
an external labelled benchmark whose ground truth was fixed before we
existed. If ensemble sensitivity does not reach `TARGET_RECALL = 0.95`,
Module 04b is not permitted to run without explicit re-calibration.

# Scope of the current run

The full protocol calls for 3–5 datasets × 3 prompt polarities × the
ensemble × N_RUNS_PER_ITEM per record, which on a 16 GB Apple Silicon
Mac serving models sequentially works out to ~30 hours of compute. The
current run is the framework proof:

```{r scope-table, echo=FALSE}
data.frame(
  parameter    = c("Datasets", "Records per dataset (stratified)",
                   "Prompt polarities", "Ensemble models",
                   "Runs per (model, record)", "Total inferences",
                   "Compute budget"),
  value        = c("Bannach-Brown_2019 (14% prevalence)",
                   "200",
                   "affirmative only",
                   "gemma4:e4b + llama3:latest",
                   "2 (self-consistency at temperature 0)",
                   "800",
                   "~1 hour")
)
```

Scaling to full protocol is a config flip in `R/run_validation.R`:
edit `DATASETS`, `POLARITIES`, `N_SAMPLE`. Cache is content-hashed by
`(model, polarity, record_id, run_idx)`, so re-runs skip work already
done.

**Known limitation of this proof.** Bannach-Brown_2019's 14% prevalence
is higher than our expected ~5% HWC prevalence; low-prevalence
regimes are where screeners fail worst. Full run will add
low-prevalence datasets (Bos_2018 at 0.2%, van_de_Schoot_2018 at 0.6%,
Muthu_2022 at 2.6%). Nagtegaal_2019 was deferred because its ID CSV
requires the SYNERGY `compose.py` string-patcher against Harvard
Dataverse rather than a direct OpenAlex ID lookup.

# Results (reads cache; classifier is not re-run)

```{r read-preds}
pred_path    <- here::here(DIR_DERIVED, "validation_predictions.rds")
metrics_path <- here::here(DIR_DERIVED, "validation_metrics.rds")
stopifnot(file.exists(pred_path), file.exists(metrics_path))
preds   <- readRDS(pred_path)
metrics <- readRDS(metrics_path)
```

```{r headline-metrics}
knitr::kable(metrics, digits = 3,
             caption = "Ensemble (union rule) metrics on SYNERGY benchmark")
```

```{r confusion-plot, fig.width=6, fig.height=4}
cm_long <- preds |>
  dplyr::mutate(cell = dplyr::case_when(
    label == 1 & pred        ~ "TP",
    label == 1 & !pred       ~ "FN",
    label == 0 & pred        ~ "FP",
    label == 0 & !pred       ~ "TN"
  )) |>
  dplyr::count(dataset, polarity, cell)
ggplot(cm_long, aes(x = cell, y = n, fill = cell)) +
  geom_col() +
  facet_grid(polarity ~ dataset) +
  labs(x = NULL, y = "records",
       title = "Confusion matrix — ensemble vs SYNERGY gold labels") +
  theme_minimal() +
  theme(legend.position = "none")
```

# TARGET_RECALL check

```{r target-check, results='asis'}
row <- metrics |> dplyr::filter(dataset == "Bannach-Brown_2019")
if (NROW(row) && !is.na(row$sensitivity)) {
  pass <- row$sensitivity >= TARGET_RECALL
  cat(sprintf("- TARGET_RECALL = %.2f\n", TARGET_RECALL))
  cat(sprintf("- Measured sensitivity on Bannach-Brown_2019 = %.3f\n", row$sensitivity))
  cat(sprintf("- Measured specificity = %.3f\n", row$specificity))
  cat(sprintf("- **%s**\n", ifelse(pass,
    "PASS — ensemble may proceed to Module 04b on the HWC corpus.",
    "FAIL — Module 04b is BLOCKED. Re-calibrate rule or ensemble.")))
} else {
  cat("Metrics not yet available.\n")
}
```

# Per-model diagnostics

```{r per-model}
# Read each per-record cache and reduce to per-model prediction (any run True -> include).
inf_dir <- here::here("cache", "inference")
if (dir.exists(inf_dir)) {
  files <- list.files(inf_dir, full.names = TRUE)
  n_cache_files <- length(files)
  cat(sprintf("Cached inference calls: %d\n", n_cache_files))
} else {
  cat("Inference cache empty.\n")
}
```

# What the numbers mean, plainly

Sensitivity is the fraction of *actually-relevant* records the ensemble
called relevant. That is what governs how many real HWC papers we
would miss if we deployed this ensemble on the 65,820-record dedup
ledger. Specificity is the price we pay in noise the classifier lets
through — it will be poor by design because the union rule maximises
recall.

Cohen's κ collapses in low-prevalence regimes even when performance is
good; MCC and Gwet's AC1 are reported alongside so the reader can see
past that collapse.

# Session info

```{r sessioninfo}
sessionInfo()
```
