---
title: "Module 04b — Screening the corpus, and how much the two models actually agreed"
author: "HWC North Africa automated evidence map"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, dpi = 300)
suppressPackageStartupMessages({
  library(here); library(dplyr); library(tibble); library(readr); library(ggplot2)
})
source(here::here("R", "config.R"))
```

```{r read}
preds <- readRDS(here::here(DIR_DERIVED, "classification_predictions.rds"))
strata <- readr::read_csv(here::here(DIR_DERIVED, "screening_strata.csv"),
                          show_col_types = FALSE)
agree <- readr::read_csv(here::here(DIR_TABLES, "04b_model_agreement.csv"),
                         show_col_types = FALSE)
contested <- readRDS(here::here(DIR_DERIVED, "contested_set.rds"))
vmet <- readr::read_csv(here::here(DIR_DERIVED, "validation_metrics.csv"),
                        show_col_types = FALSE)
.a <- function(q) agree$value[agree$quantity == q]

n_dedup    <- sum(strata$n)
n_titleonly<- strata$n[strata$stratum == "unscreened_title_only"]
n_failed   <- strata$n[strata$stratum == "unscreened_inference_failed"]
n_screened <- sum(preds$screened %in% TRUE)
n_include  <- strata$n[strata$stratum == "include"]
```

# What this module did

Every deduplicated record with an abstract was classified by both ensemble
models, `r N_RUNS_PER_ITEM` runs each at temperature `r OLLAMA_TEMPERATURE`,
against the eligibility criteria frozen in Module 01. No sampling, no active
learning, no stopping rule: a machine reads all of them, so the recall risk
that justifies early termination in human screening does not exist here.

The decision rule is `r DECISION_RULE` — include if any model in any run
includes — chosen to maximise recall at known cost to precision.

# Record arithmetic

```{r arithmetic}
knitr::kable(
  strata |> dplyr::arrange(dplyr::desc(n)) |>
    dplyr::mutate(pct = round(100 * n / n_dedup, 2)),
  caption = sprintf("Disposition of all %s deduplicated records", format(n_dedup, big.mark = ",")))
```

Of `r format(n_dedup, big.mark = ",")` deduplicated records,
`r format(n_screened, big.mark = ",")` were screened and
**`r n_include` were included — `r sprintf("%.2f%%", 100 * n_include / n_screened)`
of the screened corpus.** Two strata never reached a decision and must be
carried forward as such rather than folded into the excludes.

## The title-only stratum is the larger threat to recall

`r format(n_titleonly, big.mark = ",")` records
(`r sprintf("%.1f%%", 100 * n_titleonly / n_dedup)` of the corpus) have no
abstract in the ledger and were **not screened at all.**

This was a deliberate choice, and Module 04a is the reason: on the benchmark,
restricting to abstract-present records lifted ensemble sensitivity from
`r sprintf("%.3f", vmet$sensitivity[vmet$scope=="all" & vmet$polarity=="affirmative"])`
to
`r sprintf("%.3f", vmet$sensitivity[vmet$scope=="abstract_present" & vmet$polarity=="affirmative"])`.
A title alone does not carry enough signal for these models to apply the
criteria, and screening on titles would have injected error we could measure
but not correct.

The cost is that any relevant study sitting in those
`r format(n_titleonly, big.mark = ",")` records is invisible to this map. That
is not a rounding error, and Module 07 must not describe a thin cell as an
absence without stating it. Records missing an abstract are not missing at
random: older records, grey literature, non-English-language venues and
smaller regional journals are all over-represented among them — which is the
same material a North Africa evidence map is most at risk of under-counting.

## Inference failures

`r n_failed` records failed to return schema-valid JSON on every attempt, even
after the repair pass. They are reported as `unscreened_inference_failed`
rather than as excludes, because a record the classifier could not read is not
a record the classifier rejected. At
`r sprintf("%.2f%%", 100 * n_failed / n_screened)` of the corpus this stratum
is small enough not to move any count, but it is reported for the same reason:
a silent absorption into "exclude" is exactly the failure mode this pipeline is
supposed to make impossible.

# Self-consistency, and why the number is not impressive

```{r selfc}
sc <- agree |> dplyr::filter(grepl("^self-consistency", quantity))
knitr::kable(sc |> dplyr::mutate(value = sprintf("%.4f", value)),
             caption = "Agreement between run 1 and run 2 of the same model")
```

Both models agreed with themselves on
`r sprintf("%.4f", min(sc$value))` of records — zero flips out of
`r format(as.integer(.a("records scored by both models")), big.mark = ",")`.

**Read this as a decoding property, not as evidence of stability.** Both runs
use temperature `r OLLAMA_TEMPERATURE` and the same seed, and greedy decoding
is deterministic: identical inputs must produce identical outputs. A value of
1.00 here was guaranteed before any inference ran, and it measures nothing
about robustness. The protocol asks for the number, so it is reported, and it
is reported with this caveat every time it appears.

Module 06 measures the thing this was meant to measure, by perturbing
temperature between extraction runs so that the two runs can genuinely differ.

# Inter-model agreement

```{r agreement-table}
# Counts and coefficients share one column, so format each to its own scale
# rather than letting a single digits= setting print "115.0000".
.fmt <- function(v) ifelse(v == round(v), format(v, big.mark = ",", trim = TRUE),
                           sprintf("%.4f", v))
knitr::kable(
  agree |> dplyr::filter(!grepl("^self-consistency", quantity)) |>
    dplyr::mutate(value = .fmt(value)),
  caption = "Agreement between gemma4:e4b and llama3:latest (union within model)")
```

```{r agreement-numbers}
raw_ag  <- .a("raw inter-model agreement")
kap     <- .a("Cohen's kappa")
pab     <- .a("PABAK")
n_both  <- .a("both include")
n_gemma <- .a("only gemma includes")
n_llama <- .a("only llama includes")
n_cont  <- .a("contested set size")
```

Raw agreement is `r sprintf("%.4f", raw_ag)`, which sounds excellent and is
almost meaningless: at a positive rate of
`r sprintf("%.2f%%", 100 * n_include / n_screened)` two models that both said
"exclude" to everything would agree `r sprintf("%.2f%%", 100 * (1 - n_include / n_screened))`
of the time. The chance-corrected figure is the one to read.

**Cohen's κ = `r sprintf("%.3f", kap)`** — fair agreement at best, and far
below what would be accepted from two human screeners. PABAK is
`r sprintf("%.3f", pab)`, but PABAK is high here for the same arithmetic
reason raw agreement is: it is dominated by the agreed-exclude cell. The
honest summary is that **the two models agree overwhelmingly about what to
throw away and disagree substantially about what to keep.**

# The contested set is the finding of this module

```{r contested-numbers}
pct_cont <- 100 * n_cont / n_include
```

Of the `r n_include` records the ensemble included, only **`r n_both`
(`r sprintf("%.1f%%", 100 * n_both / n_include)`) were included by both
models.** The other `r n_cont` — **`r sprintf("%.1f%%", pct_cont)` of all
includes** — rest on a single model's vote.

```{r contested-by-model}
knitr::kable(contested |> dplyr::count(included_by, name = "n_records"),
             caption = "Which model pulled each contested record in")
```

`r n_gemma` includes come from gemma alone and `r n_llama` from llama alone, so
the ensemble is substantially gemma's judgement with a llama-sized minority
report attached.

The protocol states the interpretation in advance: *a small contested set is
evidence of a well-specified criterion, a large one is a warning about the
criteria themselves.* At `r sprintf("%.0f%%", pct_cont)` this is a large
contested set, and it is a warning.

```{r contested-by-cell, fig.width=7, fig.height=4}
cc <- contested |> dplyr::count(cell_country, name = "contested") |>
  dplyr::arrange(dplyr::desc(contested))
knitr::kable(cc, caption = "Contested records by search cell")

p <- ggplot(cc, aes(x = reorder(cell_country, contested), y = contested)) +
  geom_col(fill = "#CC6677") +
  coord_flip() +
  labs(x = NULL, y = "records included by exactly one model",
       title = "The contested set by search cell",
       subtitle = sprintf("%d of %d ensemble includes (%.0f%%) rest on one model's vote",
                          n_cont, n_include, pct_cont)) +
  theme_minimal(base_size = 11)
print(p)
dir.create(here::here(DIR_FIGURES), showWarnings = FALSE, recursive = TRUE)
ggsave(here::here(DIR_FIGURES, "04b_contested_by_cell.png"), p, width = 7, height = 4, dpi = 300)
ggsave(here::here(DIR_FIGURES, "04b_contested_by_cell.svg"), p, width = 7, height = 4)
```

# What the contested set predicted, and what happened next

A large contested set predicts a noisy include set, and that is what Module 06
stage 0 found when it checked where each included study was actually
conducted.

```{r geo-consequence}
geo <- readRDS(here::here(DIR_DERIVED, "geography_verification.rds"))
n_in_scope <- sum(geo$in_scope %in% TRUE)
knitr::kable(geo |> dplyr::count(study_country, sort = TRUE),
             caption = "Verified study location of the included records")
```

Only `r n_in_scope` of the `r NROW(geo)` includes
(`r sprintf("%.1f%%", 100 * n_in_scope / NROW(geo))`) are actually situated in
the five countries. That is the estimated screening precision on our own
corpus, and it is close to the
`r sprintf("%.2f", vmet$precision[vmet$scope=="abstract_present" & vmet$polarity=="affirmative"])`
precision Module 04a measured on the benchmark — the two independent estimates
agree, which is reassuring about the measurement even though the number itself
is poor.

**This is the union rule working as designed, not failing.** Specificity of
`r sprintf("%.3f", vmet$specificity[vmet$scope=="abstract_present" & vmet$polarity=="affirmative"])`
sounds strong, but at a true prevalence near
`r sprintf("%.1f%%", 100 * n_include / n_screened)` even strong specificity
produces more false positives than true ones. We chose that trade deliberately
to protect recall, and Module 06 stage 0 is where the cost is paid back.

# How much did we miss?

```{r recall-ci}
row <- vmet |> dplyr::filter(scope == "abstract_present", polarity == "affirmative")
ci  <- binom.test(row$tp, row$tp + row$fn)$conf.int
```

Deployed sensitivity was `r sprintf("%.3f", row$sensitivity)` on the benchmark
— but on **`r row$tp + row$fn` positive records only.** The exact 95%
confidence interval for that estimate runs from
`r sprintf("%.3f", ci[1])` to `r sprintf("%.3f", ci[2])`.

So the defensible statement is not "we missed nothing". It is that the
ensemble's sensitivity is **at least `r sprintf("%.0f%%", 100 * ci[1])` with
95% confidence**, and that the point estimate rests on a very small number of
positives. Applied to `r n_include` includes, a sensitivity as low as the
interval's lower bound would imply on the order of
`r round(n_include * (1 / ci[1] - 1))` relevant records missed among the
screened corpus — before counting the
`r format(n_titleonly, big.mark = ",")` never screened at all.

Module 07 must quote the interval, not the point estimate.

# Method parameters for the supplement

```{r params, echo=FALSE}
tibble::tibble(
  parameter = c("Ensemble", "Decision rule", "Runs per (model, record)",
                "Temperature", "Seed", "Polarity",
                "Criteria hash (xxhash64 of the prompt criteria block)",
                "Scope", "Cached inference calls"),
  value = c(paste(ENSEMBLE_MODELS, collapse = " + "), DECISION_RULE,
            as.character(N_RUNS_PER_ITEM), as.character(OLLAMA_TEMPERATURE),
            as.character(OLLAMA_SEED), "affirmative (selected on benchmark, not on this corpus)",
            "e021124a22ce6e9d", "abstract-present records only",
            format(length(list.files(here::here(DIR_CACHE_INFERENCE))), big.mark = ","))
) |> knitr::kable()
```

# Session info

```{r sessioninfo}
sessionInfo()
```
