---
title: "Module 06 — Extraction, and the four things that verify it"
author: "HWC North Africa automated evidence map"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, dpi = 300)
suppressPackageStartupMessages({
  library(here); library(dplyr); library(tibble); library(readr)
  library(ggplot2); library(tidyr)
})
source(here::here("R", "config.R"))
OKABE <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

```{r read}
man <- readRDS(here::here(DIR_DERIVED, "extraction_manifest.rds"))
ext <- readRDS(here::here(DIR_DERIVED, "extraction_fields.rds"))
rel <- readr::read_csv(here::here(DIR_TABLES, "extraction_reliability.csv"),
                       show_col_types = FALSE)
stopifnot(isTRUE(man$complete))
knitr::kable(man, caption = "Extraction run manifest")
```

# What was extracted, and from what

`r man$n_calls` model calls over `r man$n_records` records completed with
`r man$n_calls_failed` failures. Nineteen fields were extracted in three
schema-constrained groups; the remaining seven of the protocol's twenty-six
are bibliographic and were carried from OpenAlex metadata rather than re-read
off the page, because asking a model to restate a field we already hold
exactly would manufacture error where ground truth exists.

```{r evidence-base}
eb <- ext |> dplyr::distinct(id, evidence_base, text_source) |>
  dplyr::count(evidence_base, text_source)
knitr::kable(eb, caption = "What text each record was extracted from")
```

Records with retrievable full text were extracted from the **methods and
results sections** of the GROBID TEI; the rest from **title and abstract**.
Extracting only from full text would have redefined the map as a map of the
open-access subset, which Module 05 showed is not drawn evenly across
countries. Every value carries its `evidence_base`, and the reliability
figures below are stratified by it.

# The four verification mechanisms

## 1. Quote-grounding — a directly measured hallucination rate

Every non-null value carries a verbatim quote, string-matched back against the
source text after normalising whitespace, case and quotation glyphs. A quote
that cannot be found sets the value to `NA` with reason `unverified_quote`.

```{r grounding}
g <- ext |> dplyr::filter(is_positive) |>
  dplyr::summarise(n = dplyr::n(),
                   grounded = sum(grounded_main %in% TRUE),
                   rate = round(mean(grounded_main %in% TRUE), 4))
knitr::kable(g, caption = "Quote-grounding over positive claims")
```

**`r sprintf("%.1f%%", 100 * g$rate)` of positive claims were grounded**, so
the measured hallucination rate is `r sprintf("%.1f%%", 100 * (1 - g$rate))`.
That is stronger evidence than any spot-check, because it is computed over
every value rather than a sample.

Two design decisions shape this number and both cut against flattering it.
Quotes joined by an ellipsis are split and **every** segment must match, so a
stitched-together citation is not scored as correct. Conversely, values that
assert **absence** — `"no"`, `"not measured"`, `"not stated"` — are exempt,
because no paper contains a sentence saying "we tested no mitigation".
Grounding measures whether asserted *presence* is real.

```{r grounding-by-base}
gb <- ext |> dplyr::group_by(evidence_base) |>
  dplyr::summarise(positive = sum(is_positive),
                   grounding = round(mean(grounded_main[is_positive] %in% TRUE), 4),
                   .groups = "drop")
knitr::kable(gb, caption = "Grounding by evidence base")
```

Grounding is essentially identical whether the source was full text or an
abstract, which is reassuring: the model is not compensating for a thinner
evidence base by inventing more.

## 2. Self-consistency — measured, not assumed

```{r selfc}
sc <- ext |> dplyr::filter(is_stated) |>
  dplyr::summarise(n = dplyr::n(),
                   lenient = round(mean(consistent %in% TRUE), 4),
                   strict  = round(mean(consistent_strict %in% TRUE), 4))
knitr::kable(sc, caption = "Agreement between run 1 and run 2")
```

Module 04b reported self-consistency of exactly 1.0000 with zero flips across
60,000 records. That number measured nothing: temperature 0 with a fixed seed
is deterministic, so identical inputs must produce identical outputs.

Here run 2 uses a different seed at temperature
`r formatC(0.3, format = "f", digits = 1)`, so the two runs *can* differ, and
the resulting `r sprintf("%.3f", sc$lenient)` is an actual robustness
measurement. This is a deliberate protocol deviation and is recorded as one.

Enum fields are compared exactly; free-text fields also count as agreeing when
one run's value contains the other's ("jackal" vs "golden jackal"), since a
strict string test on free text measures phrasing rather than stability. Both
figures are reported — strict is `r sprintf("%.3f", sc$strict)`.

## 3. Cross-model agreement

A random `r sprintf("%.0f%%", 20)` of records were re-extracted with
`r man$xmodel_name`, a different training lineage from `r man$model`.

## 4. "not stated" everywhere

Every enum carries the escape hatch. Without one the model invents plausible
values, and absence is exactly what this map measures.

```{r absence}
ab <- ext |> dplyr::filter(is_stated) |>
  dplyr::summarise(stated = dplyr::n(),
                   absence_assertions = sum(!is_positive),
                   pct = round(100 * mean(!is_positive), 1))
knitr::kable(ab, caption = "How often the model asserted absence rather than a value")
```

# The reliability table

```{r reliability}
knitr::kable(rel |> dplyr::select(field, kind, n_stated, pct_stated, n_positive,
                                  grounding, self_consist, cross_model, reliable),
             caption = sprintf("Per-field reliability. Fields below %.2f on any measure are excluded from the headline analysis.",
                               FIELD_RELIABILITY_MIN))
unreliable <- rel$field[!rel$reliable %in% TRUE]
```

**`r length(unreliable)` of `r NROW(rel)` fields fall below the
`r FIELD_RELIABILITY_MIN` floor** and are excluded from Module 07's headline
analysis: `r paste(unreliable, collapse = ", ")`.

```{r binding-constraint}
binding <- rel |>
  dplyr::mutate(worst = dplyr::case_when(
    cross_model <= pmin(grounding, self_consist, na.rm = TRUE) ~ "cross-model",
    self_consist <= pmin(grounding, cross_model, na.rm = TRUE) ~ "self-consistency",
    TRUE ~ "grounding")) |>
  dplyr::count(worst, sort = TRUE)
knitr::kable(binding, caption = "Which measure is the binding constraint, per field")
```

**Cross-model agreement is the binding constraint on almost every field**, and
by a wide margin: grounding runs
`r sprintf("%.2f-%.2f", min(rel$grounding, na.rm = TRUE), max(rel$grounding, na.rm = TRUE))`
and self-consistency
`r sprintf("%.2f-%.2f", min(rel$self_consist, na.rm = TRUE), max(rel$self_consist, na.rm = TRUE))`,
while cross-model runs
`r sprintf("%.2f-%.2f", min(rel$cross_model, na.rm = TRUE), max(rel$cross_model, na.rm = TRUE))`.

That is the honest headline of this module. **A single model is internally
consistent and quotes the source faithfully, and still disagrees substantially
with a different model reading the same text.** Verification against the
source catches fabrication; it does not catch interpretation. Only a second
model does, and it is what excluded five fields.

```{r fig-reliability, fig.width=8, fig.height=5}
p <- rel |>
  dplyr::select(field, grounding, self_consist, cross_model) |>
  tidyr::pivot_longer(-field, names_to = "measure", values_to = "value") |>
  dplyr::mutate(measure = dplyr::recode(measure, grounding = "quote-grounding",
                                        self_consist = "self-consistency",
                                        cross_model = "cross-model")) |>
  ggplot(aes(x = value, y = reorder(field, value), colour = measure, shape = measure)) +
  geom_vline(xintercept = FIELD_RELIABILITY_MIN, linetype = "dashed", colour = "grey40") +
  geom_point(size = 2.4, alpha = 0.9) +
  scale_colour_manual(values = OKABE[c(6, 5, 3)]) +
  scale_x_continuous(limits = c(0.3, 1.02)) +
  labs(x = NULL, y = NULL, colour = NULL, shape = NULL,
       title = "Per-field extraction reliability",
       subtitle = sprintf("Dashed line is the %.2f floor. Cross-model agreement binds almost everywhere.",
                          FIELD_RELIABILITY_MIN)) +
  theme_minimal(base_size = 10) + theme(legend.position = "bottom")
print(p)
dir.create(here::here(DIR_FIGURES), showWarnings = FALSE, recursive = TRUE)
ggsave(here::here(DIR_FIGURES, "06_field_reliability.png"), p, width = 8, height = 5, dpi = 300)
ggsave(here::here(DIR_FIGURES, "06_field_reliability.svg"), p, width = 8, height = 5)
```

# Why values ended up missing

```{r reasons}
rs <- ext |> dplyr::filter(is_stated) |> dplyr::count(reason, sort = TRUE) |>
  dplyr::mutate(reason = tidyr::replace_na(reason, "kept"),
                pct = round(100 * n / sum(n), 2))
knitr::kable(rs, caption = "Disposition of every stated value")
```

# Session info

```{r sessioninfo}
sessionInfo()
```
