---
title: "Module 07 — The evidence map: where the research is, and where it is not"
author: "HWC North Africa automated evidence map"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, dpi = 300)
suppressPackageStartupMessages({
  library(here); library(dplyr); library(tibble); library(readr)
  library(ggplot2); library(tidyr)
})
source(here::here("R", "config.R"))
OKABE <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
save_fig <- function(p, name, w = 7, h = 4.5) {
  dir.create(here::here(DIR_FIGURES), showWarnings = FALSE, recursive = TRUE)
  ggsave(here::here(DIR_FIGURES, paste0(name, ".png")), p, width = w, height = h, dpi = 300)
  ggsave(here::here(DIR_FIGURES, paste0(name, ".svg")), p, width = w, height = h)
  p
}
save_tab <- function(d, name) {
  dir.create(here::here(DIR_TABLES), showWarnings = FALSE, recursive = TRUE)
  readr::write_csv(d, here::here(DIR_TABLES, paste0(name, ".csv")))
  d
}
```

```{r read-all}
tiers  <- readRDS(here::here(DIR_DERIVED, "geography_tiers.rds"))
attr   <- readRDS(here::here(DIR_DERIVED, "country_attribution.rds"))
auth   <- readRDS(here::here(DIR_DERIVED, "authorship.rds"))
ctx    <- readRDS(here::here(DIR_EXTERNAL, "country_context.rds"))
ret    <- readRDS(here::here(DIR_DERIVED, "retrieval_ledger.rds"))
conf   <- readRDS(here::here(DIR_DERIVED, "include_confidence.rds"))
contest<- readRDS(here::here(DIR_DERIVED, "contested_set.rds"))
probe  <- readRDS(here::here(DIR_DERIVED, "citation_probe.rds"))
ledg   <- readRDS(here::here(DIR_DERIVED, "dedup_ledger.rds")) |>
  dplyr::distinct(id, .keep_all = TRUE)

GREY_TYPES <- c("dissertation", "report", "preprint", "book-chapter", "book",
                "conference-paper", "conference-abstract", "other")

map <- tiers |>
  dplyr::filter(analytic) |>
  dplyr::select(id, title, tier, study_country, type) |>
  dplyr::left_join(attr |> dplyr::select(id, dplyr::all_of(COUNTRIES),
                                         n_countries, any_country,
                                         quote_grounded, dplyr::ends_with("_named")),
                   by = "id") |>
  dplyr::left_join(ledg |> dplyr::select(id, publication_year, is_oa, doi,
                                         provenance_sources, provenance_cells, n_merged),
                   by = "id") |>
  dplyr::left_join(ret |> dplyr::select(id, tei_ok), by = "id") |>
  dplyr::left_join(auth |> dplyr::select(id, n_authors, author_countries,
                                         first_incountry, last_incountry, any_incountry),
                   by = "id") |>
  dplyr::left_join(conf |> dplyr::select(id, max_confidence, min_confidence), by = "id") |>
  dplyr::mutate(
    contested = id %in% contest$id,
    stratum   = dplyr::if_else(type %in% GREY_TYPES | is.na(type), "grey", "peer-reviewed"),
    sources   = vapply(provenance_sources,
                       function(x) paste(sort(unique(unlist(x))), collapse = "+"), character(1)),
    n_cells   = lengths(provenance_cells))
n_map <- NROW(map)
```

# What this map claims, and what it cannot

This is an **automated evidence map**, not a scoping review. It makes claims
about the *distribution* of research — across countries, taxa, conflict types
and methods — not about a precisely bounded included set. That distinction is
what makes it defensible under a pipeline with no human reviewer: if
classification errors are not systematically biased by country or taxon, a
distributional claim survives them, while a boundary claim would not.

Four constraints shape everything below and are repeated at each point they
bite rather than confined to a limitations paragraph:

1. **English-indexed literature only.** `r LANGUAGE_LIMITATION`
2. **`r sum(map$tei_ok %in% TRUE)` of `r n_map` records have full text.** Fields
   requiring the methods or results section are reported over that stratum with
   its own denominator.
3. **Counts by country are not mutually exclusive.** A study covering three
   countries counts in three cells.
4. **Absence in this map is absence of *retrieved, screened, English-indexed*
   evidence.** Where that is distinguishable from a real research gap, the
   text says which it is. Where it is not, it says that too.

# From search to map: the record arithmetic

```{r arithmetic}
# The canonical flow, built and reconciled once in R/build_record_flow.R so
# that this module and the Module 08 diagram cannot disagree.
flow <- readr::read_csv(here::here(DIR_DERIVED, "record_flow.csv"), show_col_types = FALSE)
geo  <- readRDS(here::here(DIR_DERIVED, "geography_verification.rds"))
n_retrieved <- flow$n[flow$step == 1]
knitr::kable(flow |> dplyr::select(stage, n, removed, removal_reason),
             caption = "From retrieval to the analytic set")
stopifnot(flow$n[flow$step == 7] == n_map)
```

The map rests on **`r n_map` records** —
`r sprintf("%.3f%%", 100 * n_map / n_retrieved)` of the
`r format(n_retrieved, big.mark = ",")` the search retrieved. Each contraction
is documented in its own module, and the largest are not equivalent:
deduplication removed redundancy, whereas the geographic gate removed
`r sum(geo$in_scope %in% FALSE)` records the classifier had wrongly included.

Two strata were **deferred, not excluded** — `r flow$removed[flow$step == 3]`
records with no abstract and `r flow$removed[flow$step == 4]` on which
inference failed outright. They were never judged, and a flow that folded them
into "excluded" would misstate what this pipeline did.

# Geographic distribution

## Raw counts

```{r geo-counts}
country_long <- map |>
  tidyr::pivot_longer(dplyr::all_of(COUNTRIES), names_to = "country", values_to = "covered") |>
  dplyr::filter(covered %in% TRUE)

geo_tbl <- tibble::tibble(country = COUNTRIES) |>
  dplyr::left_join(country_long |> dplyr::count(country, name = "studies"), by = "country") |>
  dplyr::left_join(country_long |> dplyr::filter(n_countries == 1) |>
                     dplyr::count(country, name = "sole_focus"), by = "country") |>
  dplyr::left_join(country_long |> dplyr::filter(stratum == "peer-reviewed") |>
                     dplyr::count(country, name = "peer_reviewed"), by = "country") |>
  dplyr::left_join(country_long |> dplyr::filter(stratum == "grey") |>
                     dplyr::count(country, name = "grey"), by = "country") |>
  dplyr::left_join(country_long |> dplyr::filter(tei_ok %in% TRUE) |>
                     dplyr::count(country, name = "with_full_text"), by = "country") |>
  dplyr::mutate(dplyr::across(-country, ~ tidyr::replace_na(.x, 0L)))
knitr::kable(geo_tbl, caption = "Studies per country. Counts are not mutually exclusive; peer-reviewed and grey are reported separately and never summed into a headline.")
save_tab(geo_tbl, "07_studies_by_country")
```

```{r geo-unattributed}
n_unattr <- sum(!map$any_country %in% TRUE)
```

`r n_unattr` of the `r n_map` analytic records could not be attributed to any
of the five countries from title and abstract. They passed the in-scope gate
but name no country specifically enough to place. They are **not** distributed
across the cells above, so every count here is a floor.

## Normalised by land area and population

A raw count conflates "little research happens here" with "this is a small
country". Both denominators come from the World Bank
(land area `r unique(ctx$land_area_km2_year)`, population
`r unique(ctx$population_year)`).

```{r geo-normalised}
geo_norm <- geo_tbl |>
  dplyr::left_join(ctx |> dplyr::select(country, land_area_km2, population), by = "country") |>
  dplyr::mutate(
    per_100k_km2 = round(studies / (land_area_km2 / 1e5), 2),
    per_million_people = round(studies / (population / 1e6), 2)) |>
  dplyr::select(country, studies, land_area_km2, population,
                per_100k_km2, per_million_people) |>
  dplyr::arrange(dplyr::desc(studies))
knitr::kable(geo_norm, caption = "Research intensity normalised by land area and population",
             format.args = list(big.mark = ","))
save_tab(geo_norm, "07_studies_normalised")
```

```{r geo-gap}
gmax <- geo_norm |> dplyr::slice_max(studies, n = 1)
gmin <- geo_norm |> dplyr::slice_min(studies, n = 1)
amax <- geo_norm |> dplyr::slice_max(per_100k_km2, n = 1)
amin <- geo_norm |> dplyr::slice_min(per_100k_km2, n = 1)
```

```{r geo-gap-2}
pmax_ <- geo_norm |> dplyr::slice_max(per_million_people, n = 1)
pmin_ <- geo_norm |> dplyr::slice_min(per_million_people, n = 1)
rank_raw  <- geo_norm |> dplyr::arrange(dplyr::desc(studies)) |> dplyr::pull(country)
rank_area <- geo_norm |> dplyr::arrange(dplyr::desc(per_100k_km2)) |> dplyr::pull(country)
rank_pop  <- geo_norm |> dplyr::arrange(dplyr::desc(per_million_people)) |> dplyr::pull(country)
```

**The gap, quantified.** `r gmax$country` carries `r gmax$studies` studies and
`r gmin$country` carries `r gmin$studies` — a ratio of
`r sprintf("%.1f", gmax$studies / max(gmin$studies, 1))` to 1 in raw counts.
Per 100,000 km² the spread widens sharply, from
`r sprintf("%.2f", amax$per_100k_km2)` (`r amax$country`) to
`r sprintf("%.2f", amin$per_100k_km2)` (`r amin$country`) — a ratio of
`r sprintf("%.1f", amax$per_100k_km2 / max(amin$per_100k_km2, 0.01))` to 1,
because the countries with the fewest studies are also the largest.

**But the denominator changes the answer, and that has to be said.** Per
million people the order is different: `r pmax_$country` is highest at
`r sprintf("%.2f", pmax_$per_million_people)` and `r pmin_$country` is lowest
at `r sprintf("%.2f", pmin_$per_million_people)`.

```{r geo-rank-compare}
knitr::kable(tibble::tibble(
  rank = seq_along(rank_raw),
  `by raw count` = rank_raw,
  `per 100,000 km²` = rank_area,
  `per million people` = rank_pop),
  caption = "Country order under each denominator")
```

`r sprintf("%s ranks %s of %d on raw count, %s per unit area, and %s per million people.", gmin$country, scales::ordinal(which(rank_raw == gmin$country)), length(rank_raw), scales::ordinal(which(rank_area == gmin$country)), scales::ordinal(which(rank_pop == gmin$country)))`
A country with a small population can look research-rich per capita while
holding very few studies in absolute terms. **Neither denominator is the right
one on its own**, and this map reports all three rather than choosing the one
that tells the tidiest story. Where the three disagree, the raw count is what
a researcher planning fieldwork needs and the normalised figures are what a
funder allocating effort needs.

```{r fig-geo, fig.width=8, fig.height=4.5}
p <- geo_norm |>
  tidyr::pivot_longer(c(studies, per_100k_km2, per_million_people),
                      names_to = "measure", values_to = "value") |>
  dplyr::mutate(measure = factor(measure, c("studies", "per_100k_km2", "per_million_people"),
                                 c("studies", "per 100,000 km²", "per million people"))) |>
  ggplot(aes(x = reorder(country, value), y = value, fill = measure)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ measure, scales = "free_x") +
  coord_flip() +
  scale_fill_manual(values = OKABE[c(6, 5, 3)]) +
  labs(x = NULL, y = NULL,
       title = "Human-wildlife conflict research across North Africa",
       subtitle = sprintf("%d studies; counts are not mutually exclusive", n_map)) +
  theme_minimal(base_size = 11)
save_fig(p, "07_geographic_distribution", w = 8, h = 4.5)
```

## Is a thin cell a gap, or an artefact?

```{r geo-artefact}
art <- geo_tbl |>
  dplyr::left_join(
    ret |> dplyr::mutate(cell_country = country) |> dplyr::count(cell_country, name = "cell_records"),
    by = c("country" = "cell_country")) |>
  dplyr::left_join(
    ret |> dplyr::group_by(cell_country = country) |>
      dplyr::summarise(cell_tei_pct = round(100 * mean(tei_ok %in% TRUE), 1), .groups = "drop"),
    by = c("country" = "cell_country")) |>
  dplyr::select(country, studies, cell_records, cell_tei_pct)
knitr::kable(art, caption = "Studies found against how much the search cell returned and how much of it was readable")
```

This table is the honest reading of the one above it. A country with few
studies **and** a thin search cell **and** poor full-text retrieval has three
compounding shortfalls, and the map cannot separate them from a genuine
absence of research. Naming that country's low count as a "research gap"
without this table beside it would be the central failure mode of this study.

# Temporal

```{r temporal}
FUTURE <- as.integer(format(Sys.Date(), "%Y"))
n_future <- sum(map$publication_year > FUTURE, na.rm = TRUE)
yr <- map |> dplyr::filter(!is.na(publication_year), publication_year <= FUTURE)
by_year <- yr |> dplyr::count(publication_year, stratum, name = "n")
save_tab(by_year, "07_publications_by_year")
```

`r n_future` record`r if (n_future != 1) "s"` carr`r if (n_future == 1) "ies" else "y"`
a publication year after `r FUTURE` — an upstream metadata error, not a
prediction. Those records are excluded from the time series and reported here
rather than silently dropped.

```{r fig-temporal, fig.width=8, fig.height=4}
p <- ggplot(by_year, aes(x = publication_year, y = n, fill = stratum)) +
  geom_col() +
  scale_fill_manual(values = c("peer-reviewed" = OKABE[5], "grey" = OKABE[1])) +
  labs(x = NULL, y = "studies", fill = NULL,
       title = "Publications per year",
       subtitle = sprintf("%d studies, %d-%d; grey and peer-reviewed shown separately",
                          NROW(yr), min(yr$publication_year), max(yr$publication_year))) +
  theme_minimal(base_size = 11) + theme(legend.position = "bottom")
save_fig(p, "07_publications_by_year", w = 8, h = 4)
```

```{r temporal-recent}
half <- yr |> dplyr::mutate(era = dplyr::if_else(publication_year >= 2015, "2015+", "pre-2015")) |>
  dplyr::count(era)
```

`r half$n[half$era == "2015+"]` of `r NROW(yr)`
(`r sprintf("%.0f%%", 100 * half$n[half$era == "2015+"] / NROW(yr))`) were
published in 2015 or later. Read that as growth in *indexed, English-language*
output as much as growth in research: OpenAlex coverage of regional journals
improves sharply over the same period, and the two cannot be separated here.

```{r fig-temporal-country, fig.width=8, fig.height=4.5}
tc <- country_long |> dplyr::filter(!is.na(publication_year), publication_year <= FUTURE) |>
  dplyr::count(country, publication_year, name = "n")
p <- ggplot(tc, aes(x = publication_year, y = n, colour = country)) +
  geom_line(linewidth = 0.7) + geom_point(size = 1) +
  scale_colour_manual(values = OKABE) +
  labs(x = NULL, y = "studies", colour = NULL, title = "Publications per year by country") +
  theme_minimal(base_size = 11) + theme(legend.position = "bottom")
save_fig(p, "07_year_by_country", w = 8, h = 4.5)
```

# Language and indexing

```{r indexing}
idx <- tibble::tibble(
  quantity = c("records in the analytic set",
               "found via OpenAlex only", "found via CORE only", "found via both",
               "found by exactly one search cell",
               "open access flagged", "records in French or Arabic",
               "estimated in-scope works citation chasing would have added"),
  n = c(n_map,
        sum(map$sources == "openalex"), sum(map$sources == "core"),
        sum(map$sources == "core+openalex"),
        sum(map$n_cells == 1),
        sum(map$is_oa %in% TRUE),
        sum(ledg$language[ledg$id %in% map$id] %in% c("fr", "ar")),
        # Read from the probe's own stored estimate, never re-derived here.
        round(probe$estimate$estimate[probe$estimate$test == "strict"])))
knitr::kable(idx, caption = "How the analytic set was found")
save_tab(idx, "07_indexing")
```

**`r sum(map$sources == "core")` records came from CORE alone** and would have
been missed by an OpenAlex-only search;
**`r sum(map$n_cells == 1)` of `r n_map`
(`r sprintf("%.0f%%", 100 * mean(map$n_cells == 1))`) were returned by exactly
one search cell**, so a single term-block change would have dropped them.
That is a fragile search, and it is the strongest internal evidence that the
absences in this map are partly artefacts of how it was built.

```{r probe-estimate}
pe <- probe$estimate |> dplyr::filter(test == "strict")
```

The last row is the calibrated citation-probe estimate from Module 02b:
approximately `r round(pe$estimate)` in-scope works
(95% CI `r round(pe$ci_lo)`–`r round(pe$ci_hi)`) that backward and forward
citation chasing would have added, against a
`r n_map`-record analytic set. The interval is wide because the raw detection
was `r pe$detected` record`r if (pe$detected != 1) "s"` and the string test's
measured sensitivity is only `r sprintf("%.3f", pe$sensitivity)`. That probe
searches the citation neighbourhood of records we already found, so it bounds
what citation chasing would have added — **not** what the search missed
overall.

**Zero records are in French or Arabic.** The French cells returned 2 records
across the entire 65,820-record corpus and the Arabic cells returned none.
That is a term-coverage artefact — the non-English blocks were far narrower,
and the Latin binomials that would have caught francophone papers sat in an
English-filtered cell. Maghreb wildlife research is substantially francophone,
so **the Algerian and Moroccan cells in this map are the ones most likely to
understate real research activity**, and no count here should be read as
evidence about the francophone literature.

# Authorship equity

```{r authorship}
cov <- map |> dplyr::filter(!is.na(author_countries))
eq <- tibble::tibble(
  measure = c("first author in-country", "senior (last) author in-country",
              "any author in-country", "no in-country author"),
  n = c(sum(cov$first_incountry %in% TRUE), sum(cov$last_incountry %in% TRUE),
        sum(cov$any_incountry %in% TRUE), sum(cov$any_incountry %in% FALSE)),
  denominator = c(sum(!is.na(cov$first_incountry)), sum(!is.na(cov$last_incountry)),
                  sum(!is.na(cov$any_incountry)), sum(!is.na(cov$any_incountry)))) |>
  dplyr::mutate(pct = round(100 * n / denominator, 1))
knitr::kable(eq, caption = "In-country authorship, over records with affiliation metadata")
save_tab(eq, "07_authorship_equity")
```

`r n_map - NROW(cov)` records carry no affiliation metadata and are excluded
from these denominators rather than counted as having no in-country author —
missing data is not evidence of absence.

```{r authorship-by-country}
ac <- country_long |> dplyr::filter(!is.na(author_countries)) |>
  dplyr::group_by(country) |>
  dplyr::summarise(studies = dplyr::n(),
                   first_incountry = sum(first_incountry %in% TRUE),
                   pct = round(100 * mean(first_incountry %in% TRUE), 1),
                   .groups = "drop") |>
  dplyr::arrange(dplyr::desc(studies))
knitr::kable(ac, caption = "In-country first authorship by study country")
save_tab(ac, "07_authorship_by_country")
```

An author counts as in-country if **any** of their affiliations on that work
sits in one of the five countries, which counts a Tunisian author with a joint
French post as Tunisian. The generous reading makes the finding conservative:
if the share is low even counting joint appointments, it is genuinely low.

# Taxonomic, conflict type, and method

```{r extraction-available}
# The guard is the MANIFEST, not file existence. An earlier version of this
# chunk tested only that extraction_fields.rds existed; it passed on the
# output of a 5-record smoke run and this module published a taxonomic table
# reading "n = 4 studies" as though it described the corpus. A partial run
# must be indistinguishable from no run at all here.
man_path <- here::here(DIR_DERIVED, "extraction_manifest.rds")
ext_path <- here::here(DIR_DERIVED, "extraction_fields.rds")
rel_path <- here::here(DIR_TABLES, "extraction_reliability.csv")

man <- if (file.exists(man_path)) readRDS(man_path) else NULL
have_ext <- !is.null(man) && isTRUE(man$complete) &&
  file.exists(ext_path) && file.exists(rel_path) &&
  man$n_records >= 0.95 * n_map
if (!is.null(man)) knitr::kable(man, caption = "Module 06 extraction manifest")
```

`r if (!have_ext) sprintf("**Module 06 extraction has not completed over the analytic set, so these sections are not computed.** %s No placeholder numbers are shown, and no partial result is reported as though it described the corpus.", if (is.null(man)) "No extraction manifest exists yet." else sprintf("The manifest records a run over %d of %d analytic records with %d failed calls%s.", man$n_records, n_map, man$n_calls_failed, if (!is.na(man$limit)) sprintf(" under an explicit --limit of %d", man$limit) else "")) else ""`

```{r taxonomy, eval=have_ext}
ext <- readRDS(ext_path)
rel <- readr::read_csv(rel_path, show_col_types = FALSE)
unreliable <- rel$field[!rel$reliable %in% TRUE]
cat("Fields excluded from the headline analysis (below the",
    FIELD_RELIABILITY_MIN, "reliability floor):\n  ",
    paste(unreliable, collapse = ", "), "\n")

# FREE-TEXT NORMALISATION.
# The model returns species names as written in each paper, so one taxon
# arrives as "sea turtles", "sea turtle", "loggerhead turtle" and "loggerhead
# turtles". Counting those as four species understates every one of them and
# corrupts the taxonomic strand, which is the point of this section.
#
# Normalisation is deliberately conservative: lower-case, strip punctuation,
# collapse whitespace, and singularise a trailing "s" only on words longer
# than three characters whose ending is not itself part of the singular --
# "ss", "us", "is", "as", "os". Without the "is" guard this turned "northern
# bald ibis" into "ibi". It merges
# plurals; it does NOT merge "loggerhead turtle" into "sea turtle", because
# that is a taxonomic judgement and the model was not asked to make it. The
# raw-to-normalised mapping is printed so the merge can be audited.
.norm_species <- function(x) {
  y <- tolower(trimws(gsub("[[:punct:]]+", " ", x)))
  y <- gsub("\\s+", " ", y)
  vapply(strsplit(y, " "), function(w) {
    w <- vapply(w, function(t)
      if (nchar(t) > 3 && grepl("s$", t) && !grepl("(ss|us|is|as|os)$", t))
        sub("s$", "", t) else t, character(1))
    paste(w, collapse = " ")
  }, character(1))
}

.field_counts <- function(f, normalise = FALSE) {
  if (f %in% unreliable) return(NULL)
  d <- ext |> dplyr::filter(field == f, !is.na(value_final), value_final != "not stated")
  if (normalise) d <- d |> dplyr::mutate(value_final = .norm_species(value_final))
  d |> dplyr::count(value_final, sort = TRUE) |>
    dplyr::rename(!!f := value_final, studies = n)
}
for (f in c("taxonomic_class", "conflict_type", "marine_terrestrial", "study_design")) {
  d <- .field_counts(f)
  if (is.null(d)) { cat("\n", f, ": excluded as unreliable\n"); next }
  print(knitr::kable(head(d, 20), caption = sprintf("%s (n = %d studies with a verified value)",
                                                    f, sum(d$studies))))
  save_tab(d, paste0("07_", f))
}

sp <- .field_counts("species_common", normalise = TRUE)
if (is.null(sp)) {
  cat("\nspecies_common: excluded as unreliable\n")
} else {
  print(knitr::kable(head(sp, 25),
        caption = sprintf("species_common, plural-normalised (n = %d study-species pairs, %d distinct names)",
                          sum(sp$studies), NROW(sp))))
  save_tab(sp, "07_species_common")

  # The audit trail for the merge above.
  sp_map <- ext |>
    dplyr::filter(field == "species_common", !is.na(value_final),
                  value_final != "not stated") |>
    dplyr::mutate(normalised = .norm_species(value_final)) |>
    dplyr::count(normalised, value_final, name = "n") |>
    dplyr::group_by(normalised) |>
    dplyr::filter(dplyr::n() > 1) |>
    dplyr::ungroup() |>
    dplyr::arrange(normalised, dplyr::desc(n))
  if (NROW(sp_map)) {
    print(knitr::kable(sp_map,
          caption = sprintf("Raw forms merged by normalisation (%d names collapsed into %d)",
                            NROW(sp_map), dplyr::n_distinct(sp_map$normalised))))
    save_tab(sp_map, "07_species_normalisation_audit")
  }
}
```

```{r mitigation, eval=have_ext}
mit <- ext |> dplyr::filter(field == "mitigation_outcome_measured")
n_measured <- sum(mit$value_final %in% "yes")
cat(sprintf("Studies evaluating a mitigation with a measured outcome: %d of %d (%.1f%%)\n",
            n_measured, NROW(mit), 100 * n_measured / max(NROW(mit), 1)))
```

# Robustness

Every headline count is recomputed under defensible changes of rule. A finding
that flips is not a finding.

```{r robustness}
.recount <- function(d, label) {
  cl <- d |> tidyr::pivot_longer(dplyr::all_of(COUNTRIES), names_to = "country",
                                 values_to = "covered") |>
    dplyr::filter(covered %in% TRUE)
  tibble::tibble(variant = label, n_studies = NROW(d)) |>
    dplyr::bind_cols(as_tibble_row(setNames(
      vapply(COUNTRIES, function(c) sum(cl$country == c), integer(1)), COUNTRIES)))
}
# NOTE ON (a) AND (c). The protocol lists "union vs intersection decision
# rule" and "contested records included vs excluded" as separate variants. On
# this corpus they are the SAME set operation: a contested record is by
# definition one that exactly one model included, so dropping the contested
# set IS applying the intersection rule. They are computed once and labelled
# as one variant. Reporting them as two independent corroborations would
# double-count a single check, which is the opposite of what a robustness
# analysis is for.
variants <- dplyr::bind_rows(
  .recount(map, "headline (union rule, all)"),
  .recount(map |> dplyr::filter(!contested),
           "(a)=(c) intersection rule / contested excluded"),
  .recount(map |> dplyr::filter(max_confidence %in% "high"), "(b) high-confidence only"),
  .recount(map |> dplyr::filter(stratum == "peer-reviewed"), "(d) peer-reviewed only"),
  .recount(map |> dplyr::filter(tier != "C_asserted_no_evidence"),
           "(f) geographic tier C excluded"))
knitr::kable(variants, caption = "Studies per country under each robustness variant")
save_tab(variants, "07_robustness_variants")
```

Variants (a) and (c) are one variant, not two. A contested record is by
definition one that exactly one model included, so excluding the contested set
**is** the intersection rule. They are reported together; presenting them as
two independent checks would double-count a single test.

The intersection rule cuts the map from `r NROW(map)` studies to
`r variants$n_studies[2]` — it removes
`r sprintf("%.0f%%", 100 * (1 - variants$n_studies[2] / NROW(map)))` of the
evidence base. Variant (b) removes almost nothing
(`r NROW(map) - variants$n_studies[3]` records), because
`r sprintf("%.0f%%", 100 * mean(map$max_confidence %in% "high"))` of includes
were high-confidence. **The classifier was confidently wrong, not tentatively
wrong**, so confidence filtering is not a usable safeguard here.

```{r robustness-rank}
ranks <- variants |>
  tidyr::pivot_longer(dplyr::all_of(COUNTRIES), names_to = "country", values_to = "n") |>
  dplyr::group_by(variant) |>
  dplyr::mutate(rank = rank(-n, ties.method = "min")) |>
  dplyr::ungroup() |>
  dplyr::select(variant, country, rank) |>
  tidyr::pivot_wider(names_from = variant, values_from = rank)
knitr::kable(ranks, caption = "Rank of each country under each variant. A stable ordering is the claim this map can defend.")
save_tab(ranks, "07_robustness_ranks")

top_stable <- ranks |> dplyr::filter(dplyr::if_all(-country, ~ .x == 1))
bottom <- ranks |> dplyr::filter(dplyr::if_all(-country, ~ .x >= 4))
```

`r if (NROW(top_stable)) sprintf("**%s ranks first under every variant.**", paste(top_stable$country, collapse = " and ")) else "**No country ranks first under every variant** — the leading position is not robust and must not be reported as a finding."`
`r if (NROW(bottom)) sprintf("**%s rank%s in the bottom two under every variant**, which is the more robust half of the geographic finding: the map can defend where the evidence is thinnest more confidently than where it is densest.", paste(bottom$country, collapse = " and "), if (NROW(bottom) == 1) "s" else "") else ""`

```{r robustness-note, eval=have_ext}
cat("Variant (e), records whose fields failed quote-grounding, is computed\n",
    "over the extraction output above.\n")
```

`r if (!have_ext) "Variant (e) — with and without records whose fields failed quote-grounding — requires the Module 06 extraction output and is not yet computed." else ""`

# What this map can and cannot conclude

```{r conclusions}
n_frag <- sum(map$n_cells == 1)
n_tierC <- sum(map$tier == "C_asserted_no_evidence")
```

- **Defensible:** the *ordering* of countries by research volume, to the extent
  the robustness table above shows it stable, and the fact that
  `r n_unattr` in-scope records cannot be placed in any country at all.
- **Defensible with its qualifier:** that research is thin in the low-count
  countries *of the English-indexed literature reachable by this search*.
- **Not defensible:** any claim that a low count reflects an absence of
  research. `r n_frag` records hinge on a single search cell,
  `r n_tierC` sit in the weakest geographic tier, zero records are in French or
  Arabic, and full text was unobtainable for
  `r sum(!map$tei_ok %in% TRUE)` of `r n_map` records.

# Session info

```{r sessioninfo}
sessionInfo()
```
