---
title: "Module 07b — What these studies actually found, and where the holes are"
author: "HWC North Africa automated evidence map"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, dpi = 300)
suppressPackageStartupMessages({
  library(here); library(dplyr); library(tibble); library(readr)
  library(ggplot2); library(tidyr)
})
source(here::here("R", "config.R"))
source(here::here("R", "species_reference.R"))
OKABE <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
save_fig <- function(p, name, w = 8, h = 5) {
  dir.create(here::here(DIR_FIGURES), showWarnings = FALSE, recursive = TRUE)
  ggsave(here::here(DIR_FIGURES, paste0(name, ".png")), p, width = w, height = h, dpi = 300)
  ggsave(here::here(DIR_FIGURES, paste0(name, ".svg")), p, width = w, height = h)
  p
}
save_tab <- function(d, name) {
  dir.create(here::here(DIR_TABLES), showWarnings = FALSE, recursive = TRUE)
  readr::write_csv(d, here::here(DIR_TABLES, paste0(name, ".csv"))); d
}
```

```{r read}
man   <- readRDS(here::here(DIR_DERIVED, "extraction_manifest.rds"))
stopifnot(isTRUE(man$complete))
ext   <- readRDS(here::here(DIR_DERIVED, "extraction_fields.rds"))
rel   <- readr::read_csv(here::here(DIR_TABLES, "extraction_reliability.csv"),
                         show_col_types = FALSE)
attr_ <- readRDS(here::here(DIR_DERIVED, "country_attribution.rds"))
tiers <- readRDS(here::here(DIR_DERIVED, "geography_tiers.rds"))
ledg  <- readRDS(here::here(DIR_DERIVED, "dedup_ledger.rds")) |>
  dplyr::distinct(id, .keep_all = TRUE)
unreliable <- rel$field[!rel$reliable %in% TRUE]

# Schema-echo junk. The field spec literally reads `mitigation_type: free
# text, or "not stated"`, and the model occasionally returned the string
# "free text" as the value. These are prompt artefacts, not extractions, and
# are dropped here and counted so the defect is visible rather than silent.
JUNK <- c("free text", "n/a", "na", "none", "unknown", "not applicable")
n_junk <- sum(tolower(trimws(ext$value_final)) %in% JUNK, na.rm = TRUE)

val <- function(f) {
  ext |>
    dplyr::filter(field == f, !is.na(value_final), value_final != "not stated",
                  !(tolower(trimws(value_final)) %in% JUNK)) |>
    dplyr::select(id, value = value_final, evidence_base)
}

countries_long <- attr_ |>
  dplyr::select(id, dplyr::all_of(COUNTRIES)) |>
  tidyr::pivot_longer(-id, names_to = "country", values_to = "covered") |>
  dplyr::filter(covered %in% TRUE) |>
  dplyr::select(id, country)

ta <- ledg |>
  dplyr::filter(id %in% tiers$id[tiers$analytic]) |>
  dplyr::transmute(id, text = tolower(paste(title, dplyr::coalesce(abstract, ""))))
n_map <- NROW(ta)
```

`r n_junk` extracted values were the literal schema instruction ("free text")
rather than a value. They are dropped from everything below and reported here
because a prompt defect that silently becomes data is worse than one that is
counted.

# The map proper: conflict type by country

Marginal counts say how much research exists. They do not say **what is
studied where**, which is what a reader planning work in Libya actually needs.
Every cell below is printed, including the zeros — that is the point of the
table.

```{r conflict-matrix}
ct <- val("conflict_type") |> dplyr::inner_join(countries_long, by = "id")
ct_mat <- ct |>
  dplyr::count(value, country) |>
  tidyr::pivot_wider(names_from = country, values_from = n, values_fill = 0L) |>
  dplyr::mutate(total = rowSums(dplyr::across(dplyr::any_of(COUNTRIES)))) |>
  dplyr::arrange(dplyr::desc(total))
# Force every country column to exist even if no study populated it.
for (cn in setdiff(COUNTRIES, names(ct_mat))) ct_mat[[cn]] <- 0L
ct_mat <- ct_mat |> dplyr::select(conflict_type = value, dplyr::all_of(COUNTRIES), total)
knitr::kable(ct_mat, caption = "Conflict type by country. Counts are study-country pairs; a study spanning three countries appears in three columns.")
save_tab(ct_mat, "07b_conflict_by_country")
```

```{r conflict-zeros}
zeros <- ct_mat |>
  tidyr::pivot_longer(dplyr::all_of(COUNTRIES), names_to = "country", values_to = "n") |>
  dplyr::filter(n == 0)
n_cells <- NROW(ct_mat) * length(COUNTRIES)
```

**`r NROW(zeros)` of `r n_cells` conflict-type × country cells are empty**
(`r sprintf("%.0f%%", 100 * NROW(zeros) / n_cells)`), and the occupied cells
are highly uneven: the largest holds `r max(ct_mat$total)` study-country pairs
while `r sum(ct_mat$total <= 2)` of the `r NROW(ct_mat)` conflict types have
two or fewer across the whole region.

```{r conflict-asymmetry}
# Types documented in one country and absent in a neighbour. Restricted to
# types with enough studies overall that absence is informative rather than a
# reflection of the type being rare everywhere.
asym <- ct_mat |>
  dplyr::filter(total >= 5) |>
  tidyr::pivot_longer(dplyr::all_of(COUNTRIES), names_to = "country", values_to = "n") |>
  dplyr::group_by(conflict_type) |>
  dplyr::filter(any(n == 0) & any(n >= 3)) |>
  dplyr::summarise(
    studied_in = paste(sprintf("%s (%d)", country[n >= 3], n[n >= 3]), collapse = ", "),
    absent_in  = paste(country[n == 0], collapse = ", "),
    .groups = "drop")
knitr::kable(asym, caption = "Conflict types documented in at least one country and wholly absent in another")
save_tab(asym, "07b_conflict_asymmetry")
```

These asymmetries are the map's most actionable output **and its most
treacherous**. An ecologically implausible absence — a conflict type well
documented in Tunisia and absent in Algeria, which share the same Atlas
systems, fauna and farming — is far more likely to reflect our search than the
ecology. Each row is a research question, not a conclusion.

```{r fig-conflict-matrix, fig.width=8, fig.height=5}
p <- ct |> dplyr::count(value, country) |>
  tidyr::complete(value, country = COUNTRIES, fill = list(n = 0L)) |>
  ggplot(aes(x = country, y = reorder(value, n, sum), fill = n)) +
  geom_tile(colour = "white", linewidth = 0.6) +
  geom_text(aes(label = n, colour = n > 12), size = 3.2, show.legend = FALSE) +
  scale_fill_gradient(low = "#F7F7F7", high = OKABE[5], name = "studies") +
  scale_colour_manual(values = c("grey25", "white")) +
  labs(x = NULL, y = NULL, title = "What is studied where",
       subtitle = sprintf("%d of %d cells are empty", NROW(zeros), n_cells)) +
  theme_minimal(base_size = 11) + theme(panel.grid = element_blank())
save_fig(p, "07b_conflict_by_country", 8, 5)
```

# Species with documented conflict and zero studies

```{r species-zero}
sp_ext <- val("species_common") |> dplyr::select(id, sp = value)

ref <- SPECIES_REFERENCE |>
  dplyr::rowwise() |>
  dplyr::mutate(
    pat = species_pattern(scientific),
    ids = list(unique(c(
      ta$id[grepl(pat, ta$text)],
      sp_ext$id[grepl(pat, tolower(sp_ext$sp))])))) |>
  dplyr::ungroup() |>
  dplyr::mutate(studies = lengths(ids))

ref_tbl <- ref |>
  dplyr::select(scientific, common, conflict_context, studies) |>
  dplyr::arrange(studies, scientific)
knitr::kable(ref_tbl, caption = "The eleven taxa the protocol named as plausibly involved in regional conflict, frozen before the search ran")
save_tab(ref_tbl, "07b_species_reference_coverage")
zero_sp <- ref_tbl |> dplyr::filter(studies == 0)
thin_sp <- ref_tbl |> dplyr::filter(studies > 0, studies <= 3)
```

This reference set is **pre-specified**: it is the eleven binomials written
into the search string in Module 01, before any record was retrieved. The
search actively looked for every one of them, so a species with no studies is
one the search sought and did not find — not one we noticed was missing after
the fact.

`r if (NROW(zero_sp)) sprintf("**%d of the eleven have zero studies: %s.**", NROW(zero_sp), paste(sprintf("_%s_ (%s)", zero_sp$scientific, zero_sp$common), collapse = ", ")) else "**Every one of the eleven has at least one study**, which is itself a result: the protocol's a priori conflict taxa are all represented, so the gaps in this map are gaps of depth rather than of coverage."`

`r if (NROW(thin_sp)) sprintf("A further %d are represented by three studies or fewer: %s. At that depth a single additional study would materially change what is known.", NROW(thin_sp), paste(sprintf("_%s_ (%s, n=%d)", thin_sp$scientific, thin_sp$common, thin_sp$studies), collapse = ", ")) else ""`

```{r species-country}
# Every reference taxon must appear, including the ones with no studies at
# all: a species that drops out of the join is precisely the finding this
# table exists to report. Built by completing against the full reference set
# rather than by joining and hoping.
sp_country <- ref |>
  dplyr::select(scientific, common, ids) |>
  tidyr::unnest(ids, keep_empty = TRUE) |>
  dplyr::rename(id = ids) |>
  dplyr::left_join(countries_long, by = "id", relationship = "many-to-many") |>
  dplyr::filter(!is.na(country)) |>
  dplyr::count(scientific, common, country) |>
  tidyr::complete(tidyr::nesting(scientific, common), country = COUNTRIES,
                  fill = list(n = 0L)) |>
  tidyr::pivot_wider(names_from = country, values_from = n, values_fill = 0L) |>
  dplyr::right_join(SPECIES_REFERENCE |> dplyr::select(scientific, common),
                    by = c("scientific", "common")) |>
  dplyr::mutate(dplyr::across(dplyr::any_of(COUNTRIES), ~ tidyr::replace_na(.x, 0L)))
for (cn in setdiff(COUNTRIES, names(sp_country))) sp_country[[cn]] <- 0L
sp_country <- sp_country |>
  dplyr::select(scientific, common, dplyr::all_of(COUNTRIES)) |>
  dplyr::mutate(total = rowSums(dplyr::across(dplyr::all_of(COUNTRIES)))) |>
  dplyr::arrange(dplyr::desc(total))
stopifnot(NROW(sp_country) == NROW(SPECIES_REFERENCE))
knitr::kable(sp_country, caption = "Reference taxa by country: the species-level evidence map, zeros included")
save_tab(sp_country, "07b_species_by_country")

sp_zero_cells <- sum(sp_country |> dplyr::select(dplyr::all_of(COUNTRIES)) == 0)
```

**`r sp_zero_cells` of `r NROW(sp_country) * length(COUNTRIES)` species ×
country cells are empty.** For taxa the protocol itself flagged as regionally
conflict-prone, that is the sharpest statement of the gap this map can make.

# Did anything work? The mitigation evidence

```{r mitigation-core}
mo  <- val("mitigation_outcome_measured")
od  <- val("outcome_direction")
mt  <- val("mitigation_type")
mtd <- val("mitigation_tested")
n_measured <- sum(mo$value == "yes")
```

```{r mitigation-funnel}
funnel <- tibble::tibble(
  stage = c("studies in the map",
            "state whether a mitigation was tested",
            "a mitigation was tested",
            "state whether its outcome was measured",
            "outcome was measured",
            "report a direction of effect"),
  n = c(n_map, NROW(mtd), sum(mtd$value == "yes"), NROW(mo), n_measured,
        sum(od$value != "not measured")))
knitr::kable(funnel, caption = "The intervention-evidence funnel")
save_tab(funnel, "07b_mitigation_funnel")
```

```{r outcome-direction}
dir_tbl <- od |> dplyr::count(value, sort = TRUE) |>
  dplyr::rename(outcome_direction = value, studies = n)
knitr::kable(dir_tbl, caption = "Direction of effect, where a direction was reported")
save_tab(dir_tbl, "07b_outcome_direction")
n_reduced <- sum(od$value == "reduced conflict")
n_notmeas <- sum(od$value == "not measured")
```

Of the `r NROW(od)` studies that reported an outcome direction at all,
**`r n_notmeas` say the outcome was not measured** and `r n_reduced` report
reduced conflict, against `r sum(od$value == "increased conflict")` reporting
increased conflict and `r sum(od$value == "mixed")` mixed.

**Read the `r n_reduced` "reduced conflict" results with care.** They are not
`r n_reduced` demonstrations that something works. Nothing here distinguishes a
controlled evaluation from an author stating that a measure helped, and
positive results are far likelier to be written up than null ones. What the
map can defend is the *shape* of the funnel: interventions are described far
more often than they are evaluated.

```{r mitigation-what}
mt_tbl <- mt |>
  dplyr::mutate(value = dplyr::if_else(nchar(value) > 90,
                                       paste0(substr(value, 1, 87), "..."), value)) |>
  dplyr::arrange(value) |>
  dplyr::select(mitigation = value, evidence_base)
knitr::kable(mt_tbl, caption = "Every mitigation described in the corpus, as extracted")
save_tab(mt_tbl, "07b_mitigations_described")
```

This table is the corpus's practical content: turtle excluder devices,
acoustic pingers on nets, swordfish buoy gear replacing driftnets, barn-owl
nest boxes as rodent control, zinc phosphide and aluminium phosphide rodent
campaigns, bird netting and scarers, fenced enclosures, dog vaccination,
protected-area establishment. **`r NROW(mt)` distinct interventions across
`r n_map` studies**, most described once.

```{r mitigation-by-type}
mit_ct <- mo |> dplyr::filter(value == "yes") |>
  dplyr::inner_join(val("conflict_type") |> dplyr::select(id, ctype = value), by = "id") |>
  dplyr::count(ctype, name = "outcome_measured") |>
  dplyr::full_join(val("conflict_type") |> dplyr::count(value, name = "studies"),
                   by = c("ctype" = "value")) |>
  dplyr::mutate(dplyr::across(dplyr::everything(), ~ tidyr::replace_na(.x, 0)),
                outcome_measured = as.integer(outcome_measured),
                pct = round(100 * outcome_measured / pmax(studies, 1), 1)) |>
  dplyr::arrange(dplyr::desc(studies))
knitr::kable(mit_ct, caption = "Which conflict types have any evaluated intervention behind them")
save_tab(mit_ct, "07b_mitigation_by_conflict_type")
```

# What damage has actually been quantified

```{r damage}
de <- val("damage_estimate")
knitr::kable(de |> dplyr::select(damage_estimate = value, evidence_base),
             caption = "Every quantified damage or mortality estimate in the corpus")
save_tab(de |> dplyr::select(damage_estimate = value, evidence_base), "07b_damage_estimates")
```

`r NROW(de)` of `r n_map` studies (`r sprintf("%.0f%%", 100 * NROW(de) / n_map)`)
report a quantified damage or mortality figure. They are heterogeneous by
nature — annual turtle mortality, percentage crop loss by date cultivar,
vulture deaths across 26 countries, monetary loss in dollars — and **must not
be pooled**. No common denominator exists across them, and constructing one
would invent a comparability the literature does not have. They are listed
individually because that is the only honest presentation.

# Method mix

```{r method-mix}
if ("study_design" %in% unreliable) {
  cat("study_design fell below the reliability floor (cross-model",
      rel$cross_model[rel$field == "study_design"],
      ") and is excluded from the headline analysis.\n")
}
mm <- tibble::tibble(
  approach = c("reports a quantified damage or mortality estimate",
               "tests a mitigation",
               "measures a mitigation outcome",
               "reports stakeholder engagement"),
  studies = c(NROW(de), sum(mtd$value == "yes"), n_measured,
              sum(val("stakeholder_engagement")$value == "yes")),
  denominator = n_map) |>
  dplyr::mutate(pct = round(100 * studies / denominator, 1))
knitr::kable(mm, caption = "What kind of work this corpus contains")
save_tab(mm, "07b_method_mix")
```

`r if ("study_design" %in% unreliable) sprintf("The protocol asked for perception work versus quantified damage measurement versus intervention evaluation. The **study_design** field cannot support that split: cross-model agreement was %.3f, below the %.2f floor, so it is excluded. The table above substitutes measures built from fields that did clear the floor, and the perception-versus-measurement contrast the protocol wanted is therefore **not available** — it is a casualty of the reliability threshold, and reporting it anyway would mean using a field we have already declared unreliable.", rel$cross_model[rel$field == "study_design"], FIELD_RELIABILITY_MIN) else ""`

# What this section adds to the map

- The corpus is **overwhelmingly marine**, and its single largest conflict type
  is fisheries bycatch. Terrestrial conflict — the crop and livestock damage
  that dominates lived experience across the Maghreb — is comparatively thin,
  and **livestock depredation is nearly absent** despite jackal, wolf and
  hyaena predation being live regional issues.
- **`r sprintf("%.0f%%", 100 * NROW(zeros) / n_cells)` of conflict-type ×
  country cells are empty**, and `r sp_zero_cells` of
  `r NROW(sp_country) * length(COUNTRIES)` species × country cells.
- **Interventions are described far more often than evaluated.** Of the
  corpus, `r n_measured` studies measured a mitigation outcome.
- Whether these holes are absences of research or absences of *English-indexed,
  retrievable* research cannot be settled from inside this map. The francophone
  qualifier applies to every terrestrial Maghreb gap above.

# Session info

```{r sessioninfo}
sessionInfo()
```
