---
title: "Module 08 — Figures"
author: "HWC North Africa automated evidence map"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, dpi = 300)
suppressPackageStartupMessages({
  library(here); library(dplyr); library(tibble); library(readr)
  library(ggplot2); library(tidyr)
})
source(here::here("R", "config.R"))

# Okabe-Ito: designed for deuteranopia, protanopia and tritanopia, and it
# survives greyscale printing, which a conference poster may well get.
OKABE <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442",
           "#0072B2", "#D55E00", "#CC79A7", "#000000")

save_fig <- function(p, name, w = 7, h = 5) {
  dir.create(here::here(DIR_FIGURES), showWarnings = FALSE, recursive = TRUE)
  ggsave(here::here(DIR_FIGURES, paste0(name, ".png")), p, width = w, height = h, dpi = 300)
  ggsave(here::here(DIR_FIGURES, paste0(name, ".svg")), p, width = w, height = h)
  p
}
```

# The flow diagram

This is an **evidence-map flow diagram**, not a PRISMA-ScR diagram. It follows
PRISMA's reporting convention of showing every record and every loss, and it
deliberately does not claim the design: there was no eligibility adjudication
by a human reviewer at any stage, and two of the boxes hold records that were
never judged at all.

```{r flow-data}
flow <- readr::read_csv(here::here(DIR_DERIVED, "record_flow.csv"), show_col_types = FALSE)
knitr::kable(flow |> dplyr::select(stage, n, removed, removal_reason),
             caption = "The canonical record flow (R/build_record_flow.R)")
```

```{r fig-flow, fig.width=9, fig.height=8}
lab <- flow |>
  dplyr::mutate(
    y = rev(seq_len(dplyr::n())),
    box = sprintf("%s\nn = %s", stage, format(n, big.mark = ",", trim = TRUE)),
    side = dplyr::if_else(is.na(removed), NA_character_,
                          sprintf("−%s  %s", format(removed, big.mark = ",", trim = TRUE),
                                  removal_reason)))

p <- ggplot(lab) +
  geom_rect(aes(xmin = 0, xmax = 3.1, ymin = y - 0.36, ymax = y + 0.36),
            fill = "white", colour = "grey25", linewidth = 0.4) +
  geom_text(aes(x = 1.55, y = y, label = box), size = 3.1, lineheight = 1.05) +
  geom_segment(data = lab |> dplyr::filter(y > min(y)),
               aes(x = 1.55, xend = 1.55, y = y - 0.36, yend = y - 0.64),
               arrow = arrow(length = unit(0.16, "cm"), type = "closed"),
               linewidth = 0.35, colour = "grey25") +
  geom_text(data = lab |> dplyr::filter(!is.na(side)),
            aes(x = 3.35, y = y + 0.5, label = side),
            hjust = 0, size = 2.7, colour = "grey30") +
  scale_x_continuous(limits = c(-0.1, 8.1)) +
  scale_y_continuous(expand = expansion(add = 0.6)) +
  labs(title = "Evidence-map record flow",
       subtitle = "Not a PRISMA-ScR diagram: no human reviewer screened records at any stage") +
  theme_void(base_size = 11) +
  theme(plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(colour = "grey35", size = 9))
save_fig(p, "08_record_flow", w = 8.2, h = 8)
```

# Classifier performance

Putting the validation on the poster is the point. A conference increasingly
wary of AI-generated evidence should be able to see the measured numbers
rather than take an assurance.

```{r perf-data}
vmet <- readr::read_csv(here::here(DIR_DERIVED, "validation_metrics.csv"),
                        show_col_types = FALSE)
deployed <- vmet |> dplyr::filter(scope == "abstract_present", polarity == "affirmative")
knitr::kable(vmet |> dplyr::select(scope, polarity, n, tp, fp, fn, tn,
                                   sensitivity, specificity, precision, f1),
             digits = 3, caption = "Ensemble performance on Bannach-Brown_2019 (SYNERGY)")
```

```{r fig-performance, fig.width=8, fig.height=4.5}
perf <- vmet |>
  dplyr::select(scope, polarity, sensitivity, specificity, precision, f1) |>
  tidyr::pivot_longer(sensitivity:f1, names_to = "metric", values_to = "value") |>
  dplyr::mutate(
    scope = dplyr::recode(scope, all = "all records",
                          abstract_present = "abstract present (deployed)"),
    metric = factor(metric, c("sensitivity", "specificity", "precision", "f1")))

p <- ggplot(perf, aes(x = metric, y = value, fill = polarity)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.72) +
  geom_hline(yintercept = TARGET_RECALL, linetype = "dashed", colour = OKABE[6]) +
  annotate("text", x = 0.6, y = TARGET_RECALL + 0.03,
           label = sprintf("TARGET_RECALL = %.2f", TARGET_RECALL),
           hjust = 0, size = 2.9, colour = OKABE[6]) +
  facet_wrap(~ scope) +
  scale_fill_manual(values = OKABE[c(1, 5, 3)]) +
  scale_y_continuous(limits = c(0, 1.05), breaks = seq(0, 1, 0.25)) +
  labs(x = NULL, y = NULL, fill = "prompt polarity",
       title = "Ensemble screening performance on an external benchmark",
       subtitle = sprintf("SYNERGY Bannach-Brown_2019; %s; union decision rule",
                          paste(ENSEMBLE_MODELS, collapse = " + "))) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 20, hjust = 1))
save_fig(p, "08_classifier_performance", w = 8, h = 4.5)
```

```{r perf-ci}
ci <- binom.test(deployed$tp, deployed$tp + deployed$fn)$conf.int
```

**The figure without its interval would be misleading.** Deployed sensitivity
reads `r sprintf("%.3f", deployed$sensitivity)`, but it rests on
`r deployed$tp + deployed$fn` positive records; the exact 95% interval is
`r sprintf("%.3f", ci[1])` to `r sprintf("%.3f", ci[2])`. Any poster or slide
using this figure must carry the interval, and the caption below is written to
be used verbatim.

```{r fig-performance-ci, fig.width=7, fig.height=3.2}
cis <- vmet |>
  dplyr::rowwise() |>
  dplyr::mutate(lo = binom.test(tp, tp + fn)$conf.int[1],
                hi = binom.test(tp, tp + fn)$conf.int[2],
                label = sprintf("%s / %s",
                                dplyr::recode(scope, all = "all records",
                                              abstract_present = "abstract present"),
                                polarity)) |>
  dplyr::ungroup()

# Axis starts at 0.5, not 0. These are points with intervals rather than bars,
# so no length is being read as a magnitude, and the whole story lives between
# 0.64 and 1.00 -- a full-range axis compresses every interval into the right
# third and hides exactly the uncertainty the figure exists to show.
p <- ggplot(cis, aes(x = sensitivity, y = reorder(label, sensitivity))) +
  geom_vline(xintercept = TARGET_RECALL, linetype = "dashed", colour = OKABE[6]) +
  geom_errorbarh(aes(xmin = lo, xmax = hi), height = 0.22, colour = "grey40") +
  geom_point(aes(size = tp + fn), colour = OKABE[5]) +
  scale_x_continuous(limits = c(0.5, 1.02), breaks = seq(0.5, 1, 0.1)) +
  scale_size_continuous(range = c(2, 5), name = "positives") +
  labs(x = "sensitivity (exact 95% CI)", y = NULL,
       title = "Sensitivity with its uncertainty",
       subtitle = "The deployed configuration's 1.000 rests on 13 positive records") +
  theme_minimal(base_size = 11)
save_fig(p, "08_sensitivity_ci", w = 7, h = 3.2)
```

# Contested screening decisions

```{r fig-contested, fig.width=7, fig.height=3.6}
agree <- readr::read_csv(here::here(DIR_TABLES, "04b_model_agreement.csv"),
                         show_col_types = FALSE)
.a <- function(q) agree$value[agree$quantity == q]
cd <- tibble::tibble(
  outcome = c("both models included", "gemma4:e4b only", "llama3:latest only"),
  n = c(.a("both include"), .a("only gemma includes"), .a("only llama includes"))) |>
  dplyr::mutate(outcome = factor(outcome, outcome))

p <- ggplot(cd, aes(x = "", y = n, fill = outcome)) +
  geom_col(width = 0.55) +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3.4,
            colour = "white", fontface = "bold") +
  coord_flip() +
  scale_fill_manual(values = OKABE[c(3, 6, 2)]) +
  labs(x = NULL, y = "ensemble includes", fill = NULL,
       title = "How much of the include set rests on one model",
       subtitle = sprintf("Cohen's kappa = %.3f between the two models", .a("Cohen's kappa"))) +
  theme_minimal(base_size = 11) + theme(legend.position = "bottom")
save_fig(p, "08_contested_includes", w = 7, h = 3.6)
```

# Figure inventory

```{r inventory}
figs <- tibble::tibble(file = list.files(here::here(DIR_FIGURES), pattern = "\\.(png|svg)$")) |>
  dplyr::mutate(stem = sub("\\.(png|svg)$", "", file),
                ext = sub("^.*\\.", "", file)) |>
  dplyr::count(stem, name = "formats") |>
  dplyr::arrange(stem)
knitr::kable(figs, caption = "Every figure is written as both PNG and SVG at 300 dpi")
stopifnot(all(figs$formats == 2))
```

# Captions, written to be lifted

- **Evidence-map record flow.** Disposition of all
  `r format(flow$n[flow$step == 1], big.mark = ",")` records retrieved, from
  search to the `r flow$n[flow$step == 7]`-record analytic set. Two strata were
  deferred rather than excluded: `r format(flow$removed[flow$step == 3], big.mark = ",")`
  records had no abstract and `r flow$removed[flow$step == 4]` failed inference
  on every attempt. No human reviewer screened records at any stage.
- **Ensemble screening performance.** Sensitivity, specificity, precision and
  F1 for the two-model ensemble under three logically equivalent prompt
  phrasings, measured on the external SYNERGY benchmark
  Bannach-Brown_2019 before any application to our corpus. The deployed
  configuration (abstract present, affirmative phrasing) achieved sensitivity
  `r sprintf("%.3f", deployed$sensitivity)` (95% CI
  `r sprintf("%.2f", ci[1])`–`r sprintf("%.2f", ci[2])`, `r deployed$tp + deployed$fn`
  positives) and specificity `r sprintf("%.3f", deployed$specificity)`.
- **How much of the include set rests on one model.** Of
  `r .a("both include") + .a("only gemma includes") + .a("only llama includes")`
  records the union rule included, `r .a("both include")` were included by both
  models and the remainder by exactly one. Cohen's κ between the models was
  `r sprintf("%.3f", .a("Cohen's kappa"))`.

# Session info

```{r sessioninfo}
sessionInfo()
```
